
## Cell annotations

```{r "format CD45- marker data", include = FALSE}
cd45_so <- qread(here(params$so_dir, "cd45neg_so.qs"))

DefaultAssay(cd45_so) <- "RNA"
```

To annotate cell types, clusters are first compared with ImmGen reference data.
Cell types are then refined based on the expression of individual marker genes.
Correlation coefficients are shown for all epithelial cells, grouped based on
the original annotations from the ImmGen reference.
In this plot, cells labeled as "unassigned" were later classified as epithelial
cells based on Cdh1 and Krt8 expression.

```{r "correlation", fig.width = 6, fig.height = 4, out.width = "50%"}
typ <- "Epithelial cells"

cd45_so@meta.data %>%
  filter(cell_type == typ) %>%
  mutate(type = fct_reorder(type, r, .desc = TRUE)) %>%
  
  plot_violin(
    data_col = "r",
    cluster_col = "type",
    method = "boxplot",
    plot_colors = typ_clrs,
    color = "black",
    width = 0.75
  ) +
  base_theme +
  theme(
    aspect.ratio    = 0.9,
    legend.position = "none",
    axis.title.x    = element_blank(),
    axis.text.x     = element_text(size = txt_pt1, angle = 45, hjust = 1)
  )
```

<br>

Differentially expressed genes were identified separately for each
timepoint by comparing annotated epithelial cells with all other cells in the
CD45- datasets.
Genes were filtered to only include genes with a p value < 0.05 for all
timepoints.
Mean expression (relative) is shown for the top 50 differentially expressed
genes.

```{r "DEGs"}
deg_file <- here(params$table_dir, "epithelial_degs.tsv.gz")

# Identify DEGs
if (!file.exists(deg_file)) {
  Idents(cd45_so) <- cd45_so$cell_type
  
  epi_degs <- cd45_so %>%
    mutate_meta(
      mutate,
      tm = if_else(is.na(tm), "dual", as.character(tm))
    ) %>%
    FindConservedMarkers(
      ident.1 = "Epithelial cells",
      grouping.var = "tm"
    ) %>%
    as_tibble(rownames = "gene") %>%
    arrange(max_pval)
  
  epi_degs %>%
    write_tsv(deg_file)
}

epi_degs <- read_tsv(deg_file)

# Top DEGs
top_degs <- epi_degs %>%
  filter(max_pval < 0.05) %>%
  arrange(max_pval) %>%
  pull(gene) %>%
  head(100)
```

```{r "DEGs heatmap", fig.width = 10, fig.height = 9}
# Calculate average expression
expr <- cd45_so %>%
  mutate_meta(
    mutate,
    tm = if_else(vaccination == "dual", vaccination, as.character(tm))
  ) %>%
  AverageExpression(
    assays   = "RNA",
    slot     = "data",
    features = top_degs[1:50],
    group.by = c("cell_type", "tm")
  )

# Format heatmap data
dat <- expr$RNA %>%
  as_tibble(rownames = "gene") %>%
  pivot_longer(-gene) %>%
  separate_wider_delim(
    name,
    names = c("cell_type", "tm"),
    delim = "_"
  ) %>%
  
  group_by(tm, gene) %>%
  mutate(value = as.numeric(scale(value))) %>%
  ungroup() %>%
  mutate(
    gene      = fct_relevel(gene, rev(top_degs)),
    cell_type = fct_reorder(cell_type, value, .desc = TRUE),
    tm        = fct_relevel(tm, tms)
  )
  
# Create heatmap
dat %>%
  ggplot(aes(cell_type, gene, fill = value)) +
  geom_tile() +
  facet_wrap(~ tm, nrow = 1) +
  scale_fill_gradientn(colours = c("white", "#E7301F")) +
  base_theme +
  theme(
    legend.title = element_blank(),
    axis.text.x  = element_text(angle = 90, hjust = 1, vjust = 0.5),
    axis.title   = element_blank()
  )
```

<br>

Gene ontology terms (Biological Process) are shown for the top 100
differentially expressed genes.
The number of genes overlapping each term is shown to the right of each bar.

```{r "GO", fig.width = 10, fig.height = 5}
n_degs <- length(top_degs)

go <- top_degs %>%
  gost(
    organism = "mmusculus",
    evcodes = TRUE
  )

dat <- go$result %>%
  as_tibble() %>%
  select(-parents, -evidence_codes) %>%
  filter(
    p_value   < 0.05,
    term_size < 1000,
    source    == "GO:BP"
  ) %>%
  mutate(
    term_name = str_trunc(term_name, width = 40),
    term_name = str_c(term_name, " (", label_comma()(term_size), ")"),
    term_name = fct_inorder(term_name),
    term_name = fct_rev(term_name),
    n_lab     = str_c(intersection_size, "/", n_degs)
  ) %>%
  arrange(p_value)

dat %>%
  head(18) %>%
  ggplot(aes(-log10(p_value), term_name)) +
  geom_col(fill = "#2A8897") +
  geom_vline(xintercept = 0.5, linetype = 2) +
  geom_text(aes(label = n_lab), hjust = -0.1) +
  scale_x_continuous(expand = expansion(c(0, 0.1))) +
  base_theme +
  theme(axis.title.y = element_blank())
```

```{r "cleanup environment", include = FALSE}
rm(cd45_so)
gc()
```

---

<br>

<br>

```{r "EPITHELIAL CLUSTERING", eval = FALSE}
# * Re-cluster epithelial cells
# * Re-annotate epithelial cells using a merged reference containing Immgen
#   data and FRC subsets
# * Identify DEGs comparing the new annotations

# Cluster epithelial cells
so <- cd45_so %>%
  subset(cell_type == "Epithelial cells")

so <- so %>%
  FindVariableFeatures() %>%
  ScaleData() %>%
  RunPCA(npcs = 40) %>%
  RunUMAP(dims = 1:40) %>%
  FindNeighbors(dims = 1:40) %>%
  FindClusters(resolution = c(0.2, 0.5, 1))

# Re-annotate cell clusters
ref_gns   <- rownames(ref_frc)
ref_gns   <- ref_gns[ref_gns %in% rownames(ref_immgen)]
ref_types <- colnames(ref_frc)
ref_types <- ref_types[!ref_types %in% colnames(ref_immgen)]

ref_merge <- ref_frc[ref_gns, ref_types]
ref_merge <- bind_cols(ref_merge, ref_immgen[ref_gns, ])

so <- so %>%
  mutate_meta(
    select,
    -any_of(c("type", "r", "UMAP_1", "UMAP_2"))
  ) %>%
  clustify(
    ref_mat     = ref_merge,
    cluster_col = "RNA_snn_res.0.2",
    threshold   = 0,
    n_genes     = 2000
  )

# Identify DEGs
so <- so %>%
  mutate_meta(
    mutate,
    new_type = str_remove(type, " \\(.+$")
  )

Idents(so) <- so$new_type

epi_degs <- so %>%
  subset(tm != 2) %>%
  FindConservedMarkers(
    ident.1      = "Epithelial cells",
    grouping.var = "mouse",
    assay        = "RNA",
    slot         = "data",
    only.pos     = TRUE
  )

top_epi_degs <- epi_degs %>%
  filter(max_pval < 0.05)



# so %>%
#   plot_scatter(
#     "Pecam1",
#     size = 0.1
#   )
# 
# so %>%
#   plot_violin(
#     "Pdpn",
#     cluster_col = "type",
#     method = "boxplot",
#     top = Inf
#   )
# 
# so %>%
#   plot_scatter(
#     "RNA_snn_res.1",
#     size = 0.1
#   )
# 
# so %>%
#   plot_violin(
#     "Krt8",
#     cluster_col = "RNA_snn_res.1",
#     method = "boxplot",
#     top = Inf
#   )
# 
# Seurat::DoHeatmap(
#   so,
#   features = top_degs
# )
```

```{r "format CD45+ marker data", include = FALSE, eval = FALSE}
cd45_so <- qread(here(params$so_dir, "cd45pos_so.qs"))

DefaultAssay(cd45_so) <- "RNA"

gns <- c(
  "Ptprc",
  "Cd3d", "Cd19",
  "H2-Aa",
  "Itgax",  # DCs
  "Zbtb46", # DCs
  "Ly6c2",
  "Fcgr1",  # CD64, mon/mac
  "Itgam",  # cDC2
  "Cd14",   # mon/mac/neutrophils
  "Ly6g",   # neutrophils
  "Nkg7",   # NK cells
  "Tbx21",  # NK cells
  "Il7r"
)

lvls <- c(
  "T cells", "B cells",
  "DC", "Monocytes", "Neutrophils",
  "NK cells", "ILC"
)

# Create LEC/DC subtype boxplots
dat <- cd45_so %>%
  FetchData(c("tm", "vaccination", "cell_type", gns)) %>%
  as_tibble(rownames = ".cell_id") %>%
  pivot_longer(all_of(gns)) %>%
  mutate(
    tm        = if_else(vaccination == "dual", vaccination, as.character(tm)),
    type_tm   = str_c(cell_type, "_", tm),
    name      = fct_relevel(name, gns),
    cell_type = fct_relevel(cell_type, lvls)
  ) %>%
  arrange(cell_type) %>%
  mutate(type_tm = fct_inorder(type_tm))

# Data for n labels
n_labs <- dat %>%
  group_by(cell_type, tm, type_tm) %>%
  summarize(
    n = label_comma()(n_distinct(.cell_id)),
    .groups = "drop"
  ) %>%
  mutate(
    n_lab = str_replace(cell_type, "_", " "),
    n_lab = str_c(n_lab, " (", n, ")")
  )

n_labs <- set_names(n_labs$n_lab, n_labs$type_tm)

# Create boxplots
n_typs <- n_distinct(dat$cell_type)

dat %>%
  ggplot(aes(type_tm, value, fill = cell_type)) +
  geom_boxplot(outlier.size = 0.1, key_glyph = draw_key_point, alpha = 0.9) +
  facet_grid(name ~ tm, scales = "free") +
  guides(fill = guide_legend(override.aes = list(shape = 22, size = 4))) +
  scale_x_discrete(labels = n_labs) +
  scale_fill_manual(values = typ_clrs) +
  base_theme +
  theme(
    aspect.ratio    = 4 / n_typs,
    legend.position = "none",
    strip.clip      = "off",
    axis.title      = element_blank(),
    axis.text.x     = element_text(angle = 90, hjust = 1, vjust = 0.5)
  )
```

```{r "CD45- marker plots", eval = FALSE}
gns <- c(
  "Ptprc",
  "Prox1", "Lyve1",
  "Pdpn", "Pecam1",
  "Cd3d", "Cd19",
  "Ly6c2",
  "Nkg7",
  "Tbx21",
  "Itgax",  # DCs
  "Zbtb46", # DCs
  "Fcgr1",  # CD64
  "Itgam"
)

lvls <- c(
  "LEC", "Fibroblasts", "BEC", "Epithelial cells",
  "T cells", "B cells",
  "DC", "NK cells"
)

# Create LEC/DC subtype boxplots
dat <- cd45_so %>%
  FetchData(c("tm", "vaccination", "cell_type", gns)) %>%
  as_tibble(rownames = ".cell_id") %>%
  pivot_longer(all_of(gns)) %>%
  mutate(
    tm        = if_else(vaccination == "dual", vaccination, as.character(tm)),
    type_tm   = str_c(cell_type, "_", tm),
    name      = fct_relevel(name, gns),
    cell_type = fct_relevel(cell_type, lvls)
  ) %>%
  arrange(cell_type) %>%
  mutate(type_tm = fct_inorder(type_tm))

# Data for n labels
n_labs <- dat %>%
  group_by(cell_type, tm, type_tm) %>%
  summarize(
    n = label_comma()(n_distinct(.cell_id)),
    .groups = "drop"
  ) %>%
  mutate(
    n_lab = str_replace(cell_type, "_", " "),
    n_lab = str_c(n_lab, " (", n, ")")
  )

n_labs <- set_names(n_labs$n_lab, n_labs$type_tm)

# Create boxplots
n_typs <- n_distinct(dat$cell_type)

dat %>%
  ggplot(aes(type_tm, value, fill = cell_type)) +
  geom_boxplot(outlier.size = 0.1, key_glyph = draw_key_point, alpha = 0.9) +
  facet_grid(name ~ tm, scales = "free") +
  guides(fill = guide_legend(override.aes = list(shape = 22, size = 4))) +
  scale_x_discrete(labels = n_labs) +
  scale_fill_manual(values = typ_clrs) +
  base_theme +
  theme(
    aspect.ratio    = 4 / n_typs,
    legend.position = "none",
    strip.clip      = "off",
    axis.title      = element_blank(),
    axis.text.x     = element_text(angle = 90, hjust = 1, vjust = 0.5)
  )
```

```{r "epithelial cell marker plots OLD", fig.width = 8, fig.height = 4, eval = FALSE}
so <- load_obj(here(params$so_dir, "cd45neg_so.qs"))

gns <- c(
  "Ptprc", "Pdpn", "Pecam1",
  "Cdh1", "Krt8"
  # "Acta2", "Itga7", "Pdgfrb"
)

gns %>%
  map(~ {
    so %>%
      plot_violin(
        .x,
        cluster_col   = "cell_type",
        method        = "boxplot",
        plot_colors   = typ_clrs,
        color         = "black",
        outlier.size  = 0.1,
        outlier.alpha = 1
      ) +
      theme(axis.text.x = element_text(size = 5, color = "black"))
  }) %>%
  plot_grid(plotlist = ., nrow = 2)

rm(so)
```
