
```{r "cell type functions"}
.create_marker_boxes <- function(dat, gns, clrs = unname(typ_clrs),
                                 footer = "\n\n<br>\n\n") {
  gns %>%
    walk(~ {
      cat("\n\n####", .x, "\n\n")
      
      dat <- dat %>%
        filter(name == .x) %>%
        group_by(cell_type, mouse) %>%
        mutate(
          n = n_distinct(.cell_id),
          n_lab = str_c(cell_type, " (", label_comma()(n), ")")
        ) %>%
        group_by(cell_type) %>%
        mutate(
          med = median(value),
          q3  = boxplot.stats(value)$stats[4],
          q4  = boxplot.stats(value)$stats[5],
          max = max(value)
        ) %>%
        ungroup() %>%
        arrange(desc(med), desc(q3), desc(q4), desc(max)) %>%
        mutate(n_lab = fct_inorder(n_lab))
      
      plt <- dat %>%
        ggplot(aes(n_lab, value, fill = cell_type, color = cell_type)) +
        geom_boxplot(alpha = 0.5, outlier.size = 0.25) +
        facet_wrap(~ mouse, nrow = 1, scales = "free_x") +
        scale_fill_manual(values = clrs) +
        scale_color_manual(values = clrs) +
        labs(y = str_c(.x, " expression")) +
        base_theme +
        theme(
          aspect.ratio    = 1,
          axis.text.x     = element_text(angle = 90, hjust = 1, vjust = 0.5),
          axis.title.x    = element_blank(),
          legend.position = "none"
        )
      
      print(plt)
      cat(footer)
    })
}

.create_cluster_heats <- function(dat, x_clmn, n_min = 20, footer = "\n\n<br>\n\n<br>\n\n") {
  dat <- dat %>%
    group_by(!!sym(x_clmn), cell_type, mouse, name) %>%
    summarize(
      value = mean(value),
      n = n_distinct(.cell_id),
      .groups = "drop"
    ) %>%
    group_by(cell_type, mouse, name) %>%
    filter(row_number(desc(n)) < 10 & n > n_min) %>%
    group_by(name, mouse) %>%
    mutate(value = as.numeric(scale(value))) %>%
    group_by(cell_type, mouse, name) %>%
    filter(!all(is.na(value))) %>%
    ungroup() %>%
    mutate(
      x_var = str_c(mouse, "-", cell_type, "-", !!sym(x_clmn)),
      x_lab = if_else(n >= 1000, str_c(round(n / 1000, 1), "k"), as.character(n)),
      x_lab = str_c(!!sym(x_clmn), " (", x_lab, ")")
    ) %>%
    mutate(x_var = fct_reorder(x_var, n, .desc = TRUE))
  
  x_labs <- dat %>%
    distinct(x_var, x_lab) %>%
    pull(x_lab, x_var)
  
  # Create heatmaps
  dat$mouse %>%
    levels() %>%
    rev() %>%
    walk(~ {
      cat("\n\n####", .x, "\n\n")
      
      plt <- dat %>%
        filter(mouse == .x) %>%
        ggplot(aes(x_var, name, fill = value)) +
        geom_tile(color = ln_col) +
        facet_grid(~ cell_type, space = "free_x", scales = "free_x") +
        scale_fill_gradientn(colours = c("white", "red")) +
        scale_x_discrete(labels = x_labs) +
        labs(x = "cluster (# of cells)") +
        base_theme +
        theme(
          legend.ticks = element_blank(),
          legend.key.width = unit(7, "pt"),
          legend.title = element_blank(),
          strip.clip   = "off",
          strip.text   = element_text(size = txt_pt2, angle = 90, hjust = 0.1, vjust = 0.5),
          axis.title.y = element_blank(),
          axis.text.x  = element_text(angle = 90, vjust = 0.5, hjust = 1),
          axis.ticks.x = element_blank()
        )
      
      print(plt)
      cat(footer)
    })
}
```

```{r "format marker data", include = FALSE}
# Gene lists to plot
gns <- list(
  cd45neg_small = c(
    "Ptprc",
    "Prox1", "Lyve1",
    "Pdpn", "Pecam1",
    "Cd3d", "Cd19",
    "Ly6c2", "Nkg7", "Tbx21",
    "Itgax",  # DCs
    "Zbtb46", # DCs
    "Fcgr1",  # CD64
    "Epcam",
    "Itgam", "Cd207"
  ),
  cd45pos_small = c(
    "Ptprc",
    "Cd3d", "Cd19", "H2-Aa",
    "Itgax",  # DCs
    "Zbtb46", # DCs
    "Xcr1",
    "Sirpa",
    "Ccr7",
    "Cd207",
    "Ly6c2",
    "Fcgr1",  # CD64, mon/mac
    "Itgam",  # cDC2
    "Cd14",   # mon/mac/neutrophils
    "Ly6g",   # neutrophils
    "Nkg7",   # NK cells
    "Tbx21",  # NK cells
    "Il7r"
  ),
  lec_so = c(
    "Lyve1",
    "Marco",
    "Madcam1", "Vcam1", "Icam1",
    "Ptx3",  "Ackr4"
  ),
  dc_so = c(
    "Zbtb46", "Itgam",
    "Xcr1", "Itgax", "Sirpa",
    "Lyz2", "Ccr7",  "Siglech",
    "Cd207"
  )
)

# Cell type levels
type_lvls <- list(
  cd45neg_small = c(
    "LEC", "Fibroblasts", "BEC", "Epithelial cells",
    "T cells", "B cells",
    "DC", "NK cells"
  ),
  cd45pos_small = c(
    "T cells", "B cells",
    "DC", "Monocytes", "Neutrophils",
    "NK cells", "ILC"
  )
)

# Cell clusters to plot
clst_clmns <- list(
  cd45neg_small = "RNA_snn_res.5",
  cd45pos_small = "RNA_snn_res.5",
  lec_so     = "RNA_snn_res.10",
  dc_so      = "RNA_snn_res.5"
)

type_clmns <- list(
  cd45neg_small = "cell_type",
  cd45pos_small = "cell_type",
  lec_so     = "subtype",
  dc_so      = "subtype"
)

# Format data to plot
dat <- gns %>%
  imap(~ {
    obj <- qread(here(params$so_dir, str_c(.y, ".qs")))
    
    DefaultAssay(obj) <- "RNA"
    
    vars <- c(
      type_clmns[[.y]], "mouse",
      clst_clmns[[.y]], .x
    )
    
    res <- obj %>%
      FetchData(vars) %>%
      as_tibble(rownames = ".cell_id") %>%
      rename(cell_type = !!sym(type_clmns[[.y]])) %>%
      pivot_longer(all_of(.x)) %>%
      mutate(
        cell_type = fct_relevel(cell_type, type_lvls[[.y]]),
        mouse     = fct_relevel(mouse, mice),
        name      = fct_relevel(name, rev(.x))
      )
    
    gc()
    res
  })
```

```{r "marker heatmaps"}
heats <- dat %>%
  imap(~ {
    heat_dat <- .x %>%
      group_by(cell_type, mouse, name) %>%
      summarize(value = mean(value), .groups = "drop") %>%
      group_by(name, mouse) %>%
      mutate(value = as.numeric(scale(value))) %>%
      ungroup()
    
    # Create heatmap
    heat_dat %>%
      ggplot(aes(cell_type, name, fill = value)) +
      geom_tile() +
      facet_wrap(~ mouse, nrow = 1) +
      scale_fill_gradientn(colours = c("white", "red")) +
      base_theme +
      theme(
        legend.title = element_blank(),
        aspect.ratio = length(gns[[.y]]) / n_distinct(heat_dat$cell_type),
        axis.title   = element_blank(),
        axis.text.x  = element_text(angle = 90, hjust = 1, vjust = 0.5)
      )
  })
```

## Cell type markers

### CD45- markers {.tabset .tabset-pills}

Mean expression is shown for key marker genes for CD45- scRNA-seq datasets.

```{r "CD45- markers 1", fig.width = 10, fig.height = 4}
heats$cd45neg_small
```

```{r "CD45- markers 2", fig.width = 10, fig.height = 4, results = "asis"}
.create_marker_boxes(
  dat$cd45neg_small,
  gns$cd45neg_small,
  clrs = typ_clrs
)
```

### {.tabset .tabset-pills}

Mean expression is shown for key marker genes for each cell cluster.

```{r, fig.width = 12, fig.height = 4.5, results = "asis"}
.create_cluster_heats(
  dat$cd45neg_small,
  clst_clmns$cd45neg_small
)
```

### CD45+ markers {.tabset .tabset-pills}

Mean expression is shown for key marker genes for CD45+ scRNA-seq datasets.

```{r "CD45+ markers 1", fig.width = 10, fig.height = 4}
heats$cd45pos_small
```

```{r "CD45+ markers 2", fig.width = 10, fig.height = 4.4, results = "asis"}
.create_marker_boxes(
  dat$cd45pos_small,
  gns$cd45pos_small,
  clrs = typ_clrs
)
```

### {.tabset .tabset-pills}

Mean expression is shown for key marker genes for each cell cluster.

```{r, fig.width = 12, fig.height = 4.5, results = "asis"}
.create_cluster_heats(
  dat$cd45pos_small,
  clst_clmns$cd45pos_small
)
```

### LEC markers {.tabset .tabset-pills}

Mean expression is shown for key marker genes for LEC scRNA-seq datasets.

```{r "LEC markers 1", fig.width = 10, fig.height = 4}
heats$lec_so
```

```{r "LEC markers 2", fig.width = 10, fig.height = 4, results = "asis"}
.create_marker_boxes(
  dat$lec_so,
  gns$lec_so,
  clrs = lec_clrs
)
```

### {.tabset .tabset-pills}

Mean expression is shown for key marker genes for each cell cluster.

```{r, fig.width = 10, fig.height = 3, results = "asis"}
.create_cluster_heats(
  dat$lec_so,
  clst_clmns$lec_so,
  n_min = 0
)
```

### DC markers {.tabset .tabset-pills}

Mean expression is shown for key marker genes for DC scRNA-seq datasets.

```{r "DC markers 1", fig.width = 10, fig.height = 4}
heats$dc_so
```

```{r "DC markers 2", fig.width = 10, fig.height = 4, results = "asis"}
.create_marker_boxes(
  dat$dc_so,
  gns$dc_so,
  clrs = dc_clrs
)
```

### {.tabset .tabset-pills}

Mean expression is shown for key marker genes for each cell cluster.

```{r, fig.width = 10, fig.height = 3.8, results = "asis"}
.create_cluster_heats(
  dat$dc_so,
  clst_clmns$dc_so,
  n_min = 0,
  footer = "\n\n---\n\n<br>\n\n<br>\n\n"
)
```



```{r "format CD45+ marker data", include = FALSE, eval = FALSE}
cd45_so <- qread(here(params$so_dir, "cd45pos_so.qs"))

DefaultAssay(cd45_so) <- "RNA"

cd45pos_gns <- c(
  "Ptprc",
  "Cd3d", "Cd19",
  "H2-Aa",
  "Itgax",  # DCs
  "Zbtb46", # DCs
  "Ly6c2",
  "Fcgr1",  # CD64, mon/mac
  "Itgam",  # cDC2
  "Cd14",   # mon/mac/neutrophils
  "Ly6g",   # neutrophils
  "Nkg7",   # NK cells
  "Tbx21",  # NK cells
  "Il7r",
  "Cd207"
)

lvls <- c(
  "T cells", "B cells",
  "DC", "Monocytes", "Neutrophils",
  "NK cells", "ILC"
)

# Calculate mean expression
cd45pos_marks <- cd45_so %>%
  FetchData(c("cell_type", "mouse", cd45pos_gns)) %>%
  as_tibble(rownames = ".cell_id") %>%
  pivot_longer(all_of(cd45pos_gns)) %>%
  mutate(
    cell_type = fct_relevel(cell_type, lvls),
    mouse     = fct_relevel(mouse, mice),
    name      = fct_relevel(name, rev(cd45pos_gns))
  )

rm(cd45_so)
gc()
```

```{r "OLD format CD45- marker data", include = FALSE, eval = FALSE}
cd45_so <- qread(here(params$so_dir, "cd45neg_so.qs"))

DefaultAssay(cd45_so) <- "RNA"

cd45neg_gns <- c(
  "Ptprc",
  "Prox1", "Lyve1",
  "Pdpn", "Pecam1",
  "Cd3d", "Cd19",
  "Ly6c2",
  "Nkg7",
  "Tbx21",
  "Itgax",  # DCs
  "Zbtb46", # DCs
  "Fcgr1",  # CD64
  "Itgam",
  "Cd207"
)

sub_gns <- c(
  # "Marco", "Madcam1",
  # "Ackr4", "Icam1"
)

lvls <- c(
  "LEC", "Fibroblasts", "BEC", "Epithelial cells",
  "T cells", "B cells",
  "DC", "NK cells"
)

# Pull marker expression
cd45neg_marks <- cd45_so %>%
  FetchData(c("cell_type", "mouse", "RNA_snn_res.5", cd45neg_gns, sub_gns)) %>%
  as_tibble(rownames = ".cell_id") %>%
  pivot_longer(all_of(c(cd45neg_gns, sub_gns))) %>%
  mutate(
    cell_type = fct_relevel(cell_type, lvls),
    mouse     = fct_relevel(mouse, mice),
    name      = fct_relevel(name, rev(cd45neg_gns))
  )

rm(cd45_so)
gc()
```

```{r "format CD45- marker data", include = FALSE, eval = FALSE}
cd45_so <- qread(here(params$so_dir, "cd45neg_so.qs"))

DefaultAssay(cd45_so) <- "RNA"
```

```{r "correlation", fig.width = 6, fig.height = 4, out.width = "50%", eval = FALSE}
typ <- "Epithelial cells"

cd45_so@meta.data %>%
  filter(cell_type == typ) %>%
  mutate(type = fct_reorder(type, r, .desc = TRUE)) %>%
  
  plot_violin(
    data_col = "r",
    cluster_col = "type",
    method = "boxplot",
    plot_colors = typ_clrs,
    color = "black",
    width = 0.75
  ) +
  base_theme +
  theme(
    aspect.ratio    = 0.9,
    legend.position = "none",
    axis.title.x    = element_blank(),
    axis.text.x     = element_text(size = txt_pt1, angle = 45, hjust = 1)
  )
```

```{r "DEGs", eval = FALSE}
deg_file <- here(params$table_dir, "epithelial_degs.tsv.gz")

# Identify DEGs
if (!file.exists(deg_file)) {
  Idents(cd45_so) <- cd45_so$cell_type
  
  epi_degs <- cd45_so %>%
    mutate_meta(
      mutate,
      tm = if_else(is.na(tm), "dual", as.character(tm))
    ) %>%
    FindConservedMarkers(
      ident.1 = "Epithelial cells",
      grouping.var = "tm"
    ) %>%
    as_tibble(rownames = "gene") %>%
    arrange(max_pval)
  
  epi_degs %>%
    write_tsv(deg_file)
}

epi_degs <- read_tsv(deg_file)

# Top DEGs
top_degs <- epi_degs %>%
  filter(max_pval < 0.05) %>%
  arrange(max_pval) %>%
  pull(gene) %>%
  head(100)
```

```{r "DEGs heatmap", fig.width = 10, fig.height = 9, eval = FALSE}
# Calculate average expression
expr <- cd45_so %>%
  mutate_meta(
    mutate,
    tm = if_else(vaccination == "dual", vaccination, as.character(tm))
  ) %>%
  AverageExpression(
    assays   = "RNA",
    slot     = "data",
    features = top_degs[1:50],
    group.by = c("cell_type", "tm")
  )

# Format heatmap data
dat <- expr$RNA %>%
  as_tibble(rownames = "gene") %>%
  pivot_longer(-gene) %>%
  separate_wider_delim(
    name,
    names = c("cell_type", "tm"),
    delim = "_"
  ) %>%
  
  group_by(tm, gene) %>%
  mutate(value = as.numeric(scale(value))) %>%
  ungroup() %>%
  mutate(
    gene      = fct_relevel(gene, rev(top_degs)),
    cell_type = fct_reorder(cell_type, value, .desc = TRUE),
    tm        = fct_relevel(tm, tms)
  )
  
# Create heatmap
dat %>%
  ggplot(aes(cell_type, gene, fill = value)) +
  geom_tile() +
  facet_wrap(~ tm, nrow = 1) +
  scale_fill_gradientn(colours = c("white", "#E7301F")) +
  base_theme +
  theme(
    legend.title = element_blank(),
    axis.text.x  = element_text(angle = 90, hjust = 1, vjust = 0.5),
    axis.title   = element_blank()
  )
```

```{r "GO", fig.width = 10, fig.height = 5, eval = FALSE}
n_degs <- length(top_degs)

go <- top_degs %>%
  gost(
    organism = "mmusculus",
    evcodes = TRUE
  )

dat <- go$result %>%
  as_tibble() %>%
  select(-parents, -evidence_codes) %>%
  filter(
    p_value   < 0.05,
    term_size < 1000,
    source    == "GO:BP"
  ) %>%
  mutate(
    term_name = str_trunc(term_name, width = 40),
    term_name = str_c(term_name, " (", label_comma()(term_size), ")"),
    term_name = fct_inorder(term_name),
    term_name = fct_rev(term_name),
    n_lab     = str_c(intersection_size, "/", n_degs)
  ) %>%
  arrange(p_value)

dat %>%
  head(18) %>%
  ggplot(aes(-log10(p_value), term_name)) +
  geom_col(fill = "#2A8897") +
  geom_vline(xintercept = 0.5, linetype = 2) +
  geom_text(aes(label = n_lab), hjust = -0.1) +
  scale_x_continuous(expand = expansion(c(0, 0.1))) +
  base_theme +
  theme(axis.title.y = element_blank())
```

```{r "cleanup environment", include = FALSE, eval = FALSE}
rm(cd45_so)
gc()
```

```{r "EPITHELIAL CLUSTERING", eval = FALSE}
# * Re-cluster epithelial cells
# * Re-annotate epithelial cells using a merged reference containing Immgen
#   data and FRC subsets
# * Identify DEGs comparing the new annotations

# Cluster epithelial cells
so <- cd45_so %>%
  subset(cell_type == "Epithelial cells")

so <- so %>%
  FindVariableFeatures() %>%
  ScaleData() %>%
  RunPCA(npcs = 40) %>%
  RunUMAP(dims = 1:40) %>%
  FindNeighbors(dims = 1:40) %>%
  FindClusters(resolution = c(0.2, 0.5, 1))

# Re-annotate cell clusters
ref_gns   <- rownames(ref_frc)
ref_gns   <- ref_gns[ref_gns %in% rownames(ref_immgen)]
ref_types <- colnames(ref_frc)
ref_types <- ref_types[!ref_types %in% colnames(ref_immgen)]

ref_merge <- ref_frc[ref_gns, ref_types]
ref_merge <- bind_cols(ref_merge, ref_immgen[ref_gns, ])

so <- so %>%
  mutate_meta(
    select,
    -any_of(c("type", "r", "UMAP_1", "UMAP_2"))
  ) %>%
  clustify(
    ref_mat     = ref_merge,
    cluster_col = "RNA_snn_res.1",
    # threshold   = 0,
    n_genes     = 2000
  )

so <- so %>%
  mutate_meta(
    mutate,
    new_type = str_remove(type, " \\(.+$")
  )

# Correlation matrix
cor_mat <- so %>%
  mutate_meta(
    select,
    -any_of(c("type", "r", "UMAP_1", "UMAP_2"))
  ) %>%
  clustify(
    ref_mat     = ref_merge,
    cluster_col = "new_type",
    threshold   = 0,
    n_genes     = 2000,
    seurat_out  = FALSE
  )


# # Identify DEGs
# Idents(so) <- so$new_type
# 
# epi_degs <- so %>%
#   subset(tm != 2) %>%
#   FindConservedMarkers(
#     ident.1      = "Epithelial cells",
#     grouping.var = "mouse",
#     assay        = "RNA",
#     slot         = "data",
#     only.pos     = TRUE
#   )
# 
# top_epi_degs <- epi_degs %>%
#   filter(max_pval < 0.05)



# so %>%
#   plot_scatter(
#     "Pecam1",
#     size = 0.1
#   )
# 
# so %>%
#   plot_violin(
#     "Pdpn",
#     cluster_col = "type",
#     method = "boxplot",
#     top = Inf
#   )
# 
# so %>%
#   plot_scatter(
#     "RNA_snn_res.1",
#     size = 0.1
#   )
# 
# so %>%
#   plot_violin(
#     "Krt8",
#     cluster_col = "RNA_snn_res.1",
#     method = "boxplot",
#     top = Inf
#   )
# 
# Seurat::DoHeatmap(
#   so,
#   features = top_degs
# )
```

```{r "OLD format CD45+ marker data", include = FALSE, eval = FALSE}
cd45_so <- qread(here(params$so_dir, "cd45pos_so.qs"))

DefaultAssay(cd45_so) <- "RNA"

gns <- c(
  "Ptprc",
  "Cd3d", "Cd19",
  "H2-Aa",
  "Itgax",  # DCs
  "Zbtb46", # DCs
  "Ly6c2",
  "Fcgr1",  # CD64, mon/mac
  "Itgam",  # cDC2
  "Cd14",   # mon/mac/neutrophils
  "Ly6g",   # neutrophils
  "Nkg7",   # NK cells
  "Tbx21",  # NK cells
  "Il7r"
)

lvls <- c(
  "T cells", "B cells",
  "DC", "Monocytes", "Neutrophils",
  "NK cells", "ILC"
)

# Create LEC/DC subtype boxplots
dat <- cd45_so %>%
  FetchData(c("tm", "vaccination", "cell_type", gns)) %>%
  as_tibble(rownames = ".cell_id") %>%
  pivot_longer(all_of(gns)) %>%
  mutate(
    tm        = if_else(vaccination == "dual", vaccination, as.character(tm)),
    type_tm   = str_c(cell_type, "_", tm),
    name      = fct_relevel(name, gns),
    cell_type = fct_relevel(cell_type, lvls)
  ) %>%
  arrange(cell_type) %>%
  mutate(type_tm = fct_inorder(type_tm))

# Data for n labels
n_labs <- dat %>%
  group_by(cell_type, tm, type_tm) %>%
  summarize(
    n = label_comma()(n_distinct(.cell_id)),
    .groups = "drop"
  ) %>%
  mutate(
    n_lab = str_replace(cell_type, "_", " "),
    n_lab = str_c(n_lab, " (", n, ")")
  )

n_labs <- set_names(n_labs$n_lab, n_labs$type_tm)

# Create boxplots
n_typs <- n_distinct(dat$cell_type)

dat %>%
  ggplot(aes(type_tm, value, fill = cell_type)) +
  geom_boxplot(outlier.size = 0.1, key_glyph = draw_key_point, alpha = 0.9) +
  facet_grid(name ~ tm, scales = "free") +
  guides(fill = guide_legend(override.aes = list(shape = 22, size = 4))) +
  scale_x_discrete(labels = n_labs) +
  scale_fill_manual(values = typ_clrs) +
  base_theme +
  theme(
    aspect.ratio    = 4 / n_typs,
    legend.position = "none",
    strip.clip      = "off",
    axis.title      = element_blank(),
    axis.text.x     = element_text(angle = 90, hjust = 1, vjust = 0.5)
  )
```

```{r "CD45- marker plots", eval = FALSE}
gns <- c(
  "Ptprc",
  "Prox1", "Lyve1",
  "Pdpn", "Pecam1",
  "Cd3d", "Cd19",
  "Ly6c2",
  "Nkg7",
  "Tbx21",
  "Itgax",  # DCs
  "Zbtb46", # DCs
  "Fcgr1",  # CD64
  "Itgam"
)

lvls <- c(
  "LEC", "Fibroblasts", "BEC", "Epithelial cells",
  "T cells", "B cells",
  "DC", "NK cells"
)

# Create LEC/DC subtype boxplots
dat <- cd45_so %>%
  FetchData(c("tm", "vaccination", "cell_type", gns)) %>%
  as_tibble(rownames = ".cell_id") %>%
  pivot_longer(all_of(gns)) %>%
  mutate(
    tm        = if_else(vaccination == "dual", vaccination, as.character(tm)),
    type_tm   = str_c(cell_type, "_", tm),
    name      = fct_relevel(name, gns),
    cell_type = fct_relevel(cell_type, lvls)
  ) %>%
  arrange(cell_type) %>%
  mutate(type_tm = fct_inorder(type_tm))

# Data for n labels
n_labs <- dat %>%
  group_by(cell_type, tm, type_tm) %>%
  summarize(
    n = label_comma()(n_distinct(.cell_id)),
    .groups = "drop"
  ) %>%
  mutate(
    n_lab = str_replace(cell_type, "_", " "),
    n_lab = str_c(n_lab, " (", n, ")")
  )

n_labs <- set_names(n_labs$n_lab, n_labs$type_tm)

# Create boxplots
n_typs <- n_distinct(dat$cell_type)

dat %>%
  ggplot(aes(type_tm, value, fill = cell_type)) +
  geom_boxplot(outlier.size = 0.1, key_glyph = draw_key_point, alpha = 0.9) +
  facet_grid(name ~ tm, scales = "free") +
  guides(fill = guide_legend(override.aes = list(shape = 22, size = 4))) +
  scale_x_discrete(labels = n_labs) +
  scale_fill_manual(values = typ_clrs) +
  base_theme +
  theme(
    aspect.ratio    = 4 / n_typs,
    legend.position = "none",
    strip.clip      = "off",
    axis.title      = element_blank(),
    axis.text.x     = element_text(angle = 90, hjust = 1, vjust = 0.5)
  )
```

```{r "epithelial cell marker plots OLD", fig.width = 8, fig.height = 4, eval = FALSE}
so <- load_obj(here(params$so_dir, "cd45neg_so.qs"))

gns <- c(
  "Ptprc", "Pdpn", "Pecam1",
  "Cdh1", "Krt8"
  # "Acta2", "Itga7", "Pdgfrb"
)

gns %>%
  map(~ {
    so %>%
      plot_violin(
        .x,
        cluster_col   = "cell_type",
        method        = "boxplot",
        plot_colors   = typ_clrs,
        color         = "black",
        outlier.size  = 0.1,
        outlier.alpha = 1
      ) +
      theme(axis.text.x = element_text(size = 5, color = "black"))
  }) %>%
  plot_grid(plotlist = ., nrow = 2)

rm(so)
```
