---
title: "Antigen Tracking Figures"
author: "Ryan Sheridan"
date: "`r Sys.Date()`"
output:
  html_document:
    toc:            true
    toc_float:      true
    toc_depth:      4
    df_print:       "paged"
    code_folding:   "hide"
    self_contained: true
    highlight:      "kate"
params:
  template_dir:  "src"                                                     # template Rmarkdowns
  so_dir:        "~/Dropbox/Ryan/Projects/antigen-exchange/results/sobjs/2024-03-15"  # Seurat objects
  mod_dir:       "~/Dropbox/Ryan/Projects/antigen-exchange/results/models/2023-11-01" # ML models
  geomx_dir:     "results/geomx"
  chikv_dir:     "~/Dropbox/Ryan/Projects/morrison-lnsc/results/sobjs"     # CHIKV objects
  ref_dir:       "ref"                                                     # clustifyr references
  sample_info:   "sample_info.xlsx"
  res_dir:                                                
    value:
      exp-0: "~/Projects/antigen-tracking/results"
      exp-1: "~/Projects/tamburini-antigen-tracking/results/2022-03-11"
      exp-2: "~/Projects/tamburini-antigen-tracking/results/2022-10-28"
editor_options: 
  chunk_output_type: console
---

```{r "setup", include = FALSE}
# Default chunk options
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo    = FALSE,
  dpi     = 300
)

knitr::knit(here::here(params$template_dir, "setup.Rmd"), "")
knitr::knit(here::here(params$template_dir, "setup-ml.Rmd"), "")

# Plot parameters
min_cells <- 5
rf_typs   <- c("cLEC", "fLEC", "Collecting")
rf_tms    <- c(14, 21, 42)

ag_modules <- rf_typs %>%
  map(str_c, c("_high", "_low")) %>%
  unlist()

# Keys for naming labels
ag_clmns <- c("Ag_3wk_score", "Ag_6wk_score")

ag_labs <- set_names(
  c("Ag-score (21 day)", "Ag-score (42 day)"),
  ag_clmns
)

ag_labs_2 <- set_names(
  c("Ag-score\n21 day", "Ag-score\n42 day"),
  ag_clmns
)

# Mouse key
m_key <- c(
  d2        = "2 day",
  d14       = "14 day",
  `3wk`     = "21 day",
  `6wk`     = "42 day",
  `6wk-3wk` = "dual"
)

# Experiment to show in main figures
ag_exps <- list(LEC = "exp-1", DC = "exp-2")

# Objects to use for each cell type
ag_objs <- list(LEC = lec_so, DC = dc_so)

# LEC/DC subset colors
subtype_clrs <- list(LEC = lec_clrs, DC = dc_clrs)
```

```{r "load data"}
# Format prediction labels
# make sure the correct predictions and module scores are being used
# once the models are finalized, this information can be saved in the meta.data
lec_so <- lec_so %>%
  AddMetaData(rf_preds, col.name = str_c("rf_", colnames(rf_preds)))

lec_so <- lec_so %>%
  mutate_meta(
    mutate,
    pred_grp = ifelse(
      rf_pred == "high" & !rf_correct,
      "Ag-competent", str_c("Ag-", !!sym(rf_dat_clmn))
    ),
    rf_pred = str_c("Ag-", rf_pred),
    mouse   = fct_relevel(mouse, mice),
    
    # Label data used for training models
    # d2 mouse was excluded
    training_data = subtype %in% rf_mods_best$subtype & is.na(rf_pred),
    training_data = ifelse(tm == 2, FALSE, training_data)
  )

# Add module scores
rf_feats %>%
  iwalk(~ {
    clmn_nms <- str_c(.y, "_", names(.x))

    clmn_nms <- set_names(
      str_c(clmn_nms, seq_along(clmn_nms)),
      clmn_nms
    )

    lec_so <<- lec_so %>%
      mutate_meta(dplyr::select, -any_of(names(clmn_nms))) %>%
      AddModuleScore(
        features = .x,
        name     = names(clmn_nms)
      ) %>%
      mutate_meta(dplyr::rename, !!!syms(clmn_nms))
  })
```

---

<br>

```{r "CD45- heatmaps", fig.width = 30, fig.height = 20, eval = FALSE}
# Function to create marker heatmaps
create_marker_heatmap <- function(so_in, markers) {
  
  gns <- markers %>%
    pull(gene) %>%
    as.character()
  
  clmns <- c("sample", "mouse", "subtype", "cell_type", gns)
  
  dat <- so_in %>%
    FetchData(clmns) %>%
    as_tibble(rownames = ".cell_id") %>%
    pivot_longer(any_of(gns), names_to = "gene") %>%
    left_join(markers, by = "gene") %>%
    
    group_by(subtype, mouse, gene, label) %>%
    summarize(
      pct_expr = sum(value > 0) / n(),
      pct_expr = as.character(round(pct_expr, 2)),
      value    = mean(value),
      .groups  = "drop"
    ) %>%
    group_by(gene, mouse) %>%
    mutate(value = as.numeric(scale(value))) %>%
    ungroup() %>%
    mutate(
      subtype = fct_relevel(subtype, levels(markers$annotation)),
      gene    = fct_relevel(gene, rev(levels(markers$gene)))
    )
  
  res <- dat %>%
    ggplot(aes(subtype, gene, fill = value)) +
    geom_tile() +
    geom_text(aes(label = pct_expr), size = 6 / .pt) +
    facet_grid(label ~ mouse, scales = "free_y", space = "free_y") +
    scale_fill_gradientn(colours = c("lightblue", "white", "red")) +
    base_theme +
    theme(
      strip.clip   = "off",
      strip.text.y = element_text(angle = 0, hjust = 0),
      axis.title   = element_blank(),
      axis.text.x  = element_text(angle = 45, hjust = 1)
    )
  
  res
}

# Load data
rm(lec_so, dc_so)

cd45neg_so <- qread(here(params$so_dir, "cd45neg_so.qs"))
cd45pos_so <- qread(here(params$so_dir, "cd45pos_so.qs"))

# Load cell type markers
xlsx_path <- here(params$ref_dir, "cell-type-markers.xlsx")
markers   <- xlsx::read.xlsx(xlsx_path, sheetName = "markers")

markers <- markers %>%
  mutate(
    cell_type = fct_inorder(cell_type),
    subtype   = fct_inorder(subtype)
  ) %>%
  group_by(gene) %>%
  summarize(
    label      = str_c(cell_type, collapse = "/"),
    cell_type  = dplyr::first(cell_type),
    annotation = dplyr::first(subtype),
    .groups    = "drop"
  ) %>%
  arrange(cell_type, annotation) %>%
  mutate(
    gene  = fct_inorder(gene),
    label = fct_inorder(label)
  )

# CD45- heatmaps
create_marker_heatmap(cd45neg_so, markers)
```

```{r "CD45+ heatmaps", fig.width = 25, fig.height = 20, eval = FALSE}
create_marker_heatmap(cd45pos_so, markers)

# Load data for other figures
lec_so <- qread(here(params$so_dir, "lec_so.qs"))
dc_so  <- qread(here(params$so_dir, "dc_so.qs"))

rm(cd45neg_so, cd45pos_so)
```

```{r "MARKER PLOTS", eval = FALSE}
cd45neg_so %>%
  plot_scatter(
    data_col = "cell_type",
    x = "hUMAP_1",
    y = "hUMAP_2",
    group_col = "mouse",
    size = 0.1
  )

cd45pos_so %>%
  plot_violin(
    "Fcer1g",
    "cell_type",
    "mouse",
    method = "boxplot"
  )

cd45pos_so %>%
  plot_scatter(
    data_col = "Ceacam8",
    x = "hUMAP_1",
    y = "hUMAP_2",
    group_col = "mouse",
    size = 0.1,
    plot_colors = c("lightblue", "white", "red")
  )

cd45pos_so %>%
  plot_scatter(
    data_col = "Zbtb46",
    x = "hUMAP_1",
    y = "hUMAP_2",
    group_col = "mouse",
    size = 0.1,
    plot_colors = c("lightblue", "white", "red")
  )
```

## Figure 1

Antigen persists in discrete cell populations within the lymph node

A.  Experimental design
B.  Mean Ag-score is shown for each cell type for 2, 14, 21, and 42 days post
    vaccination.
C.  UMAP projection shows LEC subsets.
D.  UMAP projections show Ag-scores for LEC subsets for each timepoint.
E.  Mean Ag-score is shown for LEC subsets for each timepoint.
F.  Ag-scores are shown for each timepoint for each LEC subset.
G.  UMAP projection shows DC subsets.
H.  UMAP projections show Ag-scores for DC subsets for each timepoint.
I.  Mean Ag-score is shown for DC subsets for each timepoint.
J.  Ag-scores are shown for each timepoint for each DC subset.

<u>Notes</u>

* Only experiment 1 is shown for the LEC 21 and 42 day timepoints (including
  for the UMAPs). Experiment 2 data include very few LECs for the 21 day
  timepoint.
* Experiment 2 is shown for the DC 21 and 42 day timepoints. Experiment 1 data
  are lower quality (sparse signal, mostly zeros).
* Subsets are only shown if there are >= `r min_cells` cells for each timepoint
* CD45+ and CD45- cell types are only shown for the CD45+ and CD45- datasets,
  respectively.

```{r "type Ag heat", fig.width = 16, fig.height = 5}
# Format Ag data for broad cell types
# exclude CD45+ cell types from CD45- datasets and vice versa
cd45neg_typs <- c("LEC", "BEC", "Fibroblasts", "Epithelial cells")

typ_ag_dat <- all_meta %>%
  filter(
    vaccination == "single",
    cell_type != "unassigned"
  ) %>%
  group_by(cell_sort) %>%
  mutate(
    tm        = fct_relevel(as.character(tm), tms),
    cell_type = fct_infreq(cell_type),
    cell_type = fct_relevel(cell_type, c("LEC", "Fibroblasts", "DC"))
  ) %>%
  ungroup() %>%
  filter(
    (cell_sort == "CD45neg" & cell_type %in% cd45neg_typs) |
      (cell_sort == "CD45pos" & !cell_type %in% cd45neg_typs)
  )

# Format data for heatmap
# exclude unassigned cells since these likely include multiple cell types
heat_dat <- typ_ag_dat %>%
  group_by(tm, cell_type) %>%
  summarize(
    Ag_score = mean(Ag_score), .groups = "drop",
    n        = n()
  ) %>%
  group_by(cell_type) %>%
  mutate(cell_type = str_c(cell_type, "\nn = ", label_comma()(sum(n)))) %>%
  ungroup() %>%
  mutate(
    tm      = fct_relevel(as.character(tm), rev(as.character(tms))),
    cell_type = fct_reorder(cell_type, Ag_score, mean, .desc = TRUE)
  )

# Create heatmap summarizing Ag scores
typ_heat <- heat_dat %>%
  ggplot(aes(cell_type, tm, fill = Ag_score)) +
  geom_tile() +
  scale_fill_gradientn(colours = c("white", "#D55E00"), na.value = "black") +
  guides(fill = guide_colorbar(
    title.position = "top", ticks = FALSE, title = "Ag-score",
    barwidth = unit(150, "pt"), barheight = unit(10, "pt")
  )) +
  labs(y = "days post vaccination") +
  base_theme +
  theme(
    plot.margin  = margin(0, 15, 0, 15),
    legend.position = "top",
    legend.justification = 0,
    legend.text  = element_text(size = txt_pt1),
    aspect.ratio = n_distinct(heat_dat$tm) / n_distinct(heat_dat$cell_type) * 0.9,
    axis.title.x = element_blank(),
    axis.text.x  = element_text(angle = 45, hjust = 1),
  )
```

```{r "subtype Ag UMAPs"}
# Data for UMAPs
# show exp1 for LECs and exp2 for DCs
fig1_dat <- ag_objs %>%
  imap(~ {
    .x@meta.data %>%
      as_tibble(rownames = ".cell_id") %>%
      filter(
        exp %in% c("exp-0", ag_exps[[.y]]),
        vaccination == "single"
      ) %>%
      mutate(tm_lab = factor(as.character(tm), tms))    
  })

# Subtype UMAPs
fig1_typ_u <- fig1_dat %>%
  imap(~ {
    typ_lvls <- .x %>%
      pull(subtype) %>%
      unique() %>%
      as.character()
    
    res <- .x %>%
      plot_scatter(
        "subtype", "hUMAP_1", "hUMAP_2",
        size         = 0.1,
        plot_colors  = c(lec_clrs, dc_clrs),
        plot_lvls    = rev(typ_lvls),
        label_params = list(size = ttl_pt1),
        panel_nrow   = 2
      ) +
      guides(color = guide_legend(ncol = 1, override.aes = list(size = 4))) +
      umap_theme +
      theme(
        aspect.ratio = 0.9,
        legend.position = "right",
        legend.title = element_blank(),
        legend.text  = element_text(size = txt_pt2)
      )
    
    res
  })

# Ag-score UMAPs
u_clrs <- list(
  LEC = c("lightblue", "white", "#D7301F"),
  DC  = c("lightblue", "white", "#6A51A3", "#6A51A3")
  # DC  = c("white", "#6A51A3")
)

fig1_ag_u <- fig1_dat %>%
  imap(~ {
    .x %>%
      filter(tm_lab %in% as.character(tms)) %>%
      plot_scatter(
        "Ag_score", "hUMAP_1", "hUMAP_2",
        plot_colors  = u_clrs[[.y]],
        label_params = list(size = ttl_pt1),
        group_col    = "tm_lab",
        panel_nrow   = 1,
        outline      = TRUE,
        size         = 0.2,
        stroke       = 0.7
      ) +
      facet_wrap(
        ~ tm_lab, nrow = 1,
        labeller = as_labeller(function(x) str_c("day ", x))
      ) +
      guides(fill = guide_colorbar(
        title.position = "right", ticks = FALSE, title = "Ag-score",
        barheight = unit(120, "pt"), barwidth = unit(10, "pt")
      )) +
      umap_theme +
      theme(
        aspect.ratio         = 0.9,
        panel.border         = element_rect(color = ln_col, linewidth = ln_pt, fill = NA),
        legend.position      = "right",
        legend.justification = 0,
        legend.title         = element_text(angle = 270, hjust = 0.5),
        strip.text.y         = element_blank()
      )
  })
```

```{r "subtype Ag boxplots", fig.width = 8, fig.height = 2}
# Format data for boxplots
ag_bx_dat <- fig1_dat %>%
  map(~ {
    .x %>%
      filter(subtype != "unassigned") %>%
      mutate(mouse = as.character(mouse)) %>%
      group_by(subtype) %>%
      filter(all(table(mouse) >= min_cells)) %>%
      ungroup() %>%
      mutate(subtype = fct_reorder(subtype, Ag_score, mean, .desc = TRUE))
  })
  
# Create boxplots
ag_bxs <- ag_bx_dat %>%
  imap(~ create_ag_boxes(.x, clrs = subtype_clrs[[.y]]))

# Pull subsets used for main figures
ag_plt_typs <- ag_bx_dat %>%
  map(~ levels(.x$subtype))
```

```{r "subtype Ag heat"}
# Create heatmaps
heat_clrs <- c(LEC = "#D7301F", DC = "#6A51A3")

ag_heats <- ag_bx_dat %>%
  imap(~ {
    dat <- .x %>%
      group_by(mouse, tm, subtype) %>%
      summarize(Ag_score = mean(Ag_score), .groups = "drop") %>%
      mutate(
        subtype = fct_reorder(subtype, Ag_score, mean),
        tm      = fct_relevel(as.character(tm), as.character(tms)),
        ttl     = "Ag-score"
      )

    rat <- n_distinct(dat$subtype) / n_distinct(dat$tm)

    dat %>%
      ggplot(aes(tm, subtype, fill = Ag_score)) +
      geom_tile() +
      facet_wrap(~ ttl) +
      labs(x = "days post vaccination") +
      scale_fill_gradientn(colours = c("lightblue", "white", heat_clrs[[.y]])) +
      guides(fill = guide_colorbar(
        ticks = FALSE, title = "Ag-score",
        barheight = unit(105, "pt"), barwidth = unit(10, "pt")
      )) +
      base_theme +
      theme(
        legend.title = element_blank(),
        aspect.ratio = rat,
        axis.title.y = element_blank(),
        plot.margin  = margin(0, 15, 0, 15),
        plot.title   = element_text(hjust = 0.5, size = ttl_pt2)
      )
  })
```

```{r "Fig 1", fig.width = 15.5, fig.height = 15.5}
# Create mock panel A
mock_a <- ggplot() +
  geom_blank() +
  theme(panel.background = element_blank())

# Create final figure
fig1 <- (mock_a + typ_heat) /
  (fig1_typ_u$LEC + fig1_ag_u$LEC) /
  (ag_heats[[1]] + ag_bxs$LEC) /
  (fig1_typ_u$DC + fig1_ag_u$DC) /
  (ag_heats[[2]] + ag_bxs$DC) +
  plot_layout(heights = c(0.7, 1, 0.7, 1, 0.7), tag_level = "keep") +
  plot_annotation(tag_levels = "A") &
  theme(
    plot.tag = element_text(size = ttl_pt2 * 1.5, face = "bold"),
    plot.tag.position = c(0, 1)
  )

fig1
```



```{r "TEST NORMALIZATION", eval = FALSE}
# Normalizing by filtered counts and subtracting T/B signal
# shows no difference between single and dual for 6wk
# all other variations of normalization show significant differences
# between single and dual

cd45neg_so <- qread(here(params$so_dir, "cd45neg_so.qs"))

clmns <- c(
  "ovalbumin", "ABPS2", "ABPS4",
  "ovalbumin_score", "Ag_3wk_score", "Ag_6wk_score"
)

ag_dat <- cd45neg_so %>%
  mutate_meta(dplyr::select, -any_of(clmns)) %>%
  calc_ag_score(
    clmns        = c("ovalbumin", "ABPS2", "ABPS4"),
    # control_types = NULL,
    norm_factors = norm_factors,
    slot         = "data",
    operation    = `-`,
    final_trans  = (function(x) x)
    # final_trans  = (function(x) log10(x + 1))
  )

clr_dat <- cd45neg_so@meta.data %>%
  dplyr::select(-any_of(c("ovalbumin_score", "Ag_3wk_score", "Ag_6wk_score"))) %>%
  rownames_to_column("cell_id") %>%
  # filter(exp != "exp-0") %>%
  pivot_longer(all_of(c("ovalbumin", "ABPS2", "ABPS4"))) %>%
  group_by(directory) %>%
  mutate(value = as.numeric(clr(na.omit(value) + 1))) %>%
  ungroup() %>%
  mutate(name = str_c(name, "_score")) %>%
  pivot_wider()

clr_dat <- clr_dat %>%
  mutate(
    Ag_score = case_when(
      exp == "exp-0" ~ ovalbumin_score,
      exp == "exp-1" & tm == 21 ~ ABPS4_score,
      exp == "exp-1" & tm == 42 ~ ABPS2_score,
      exp == "exp-2" & tm == 21 ~ ABPS2_score,
      exp == "exp-2" & tm == 42 ~ ABPS4_score
    ),
    Ag_score_3 = case_when(
      exp == "exp-1" & (tm == 21 | vaccination == "dual") ~ ABPS4_score,
      exp == "exp-2" & (tm == 21 | vaccination == "dual") ~ ABPS2_score
    ),
    Ag_score_6 = case_when(
      exp == "exp-1" & (tm == 42 | vaccination == "dual") ~ ABPS2_score,
      exp == "exp-2" & (tm == 42 | vaccination == "dual") ~ ABPS4_score
    ),
    mouse = fct_relevel(mouse, mice)
  ) %>%
  column_to_rownames("cell_id")

plt_dat <- lec_so %>%
  AddMetaData(clr_dat[c("Ag_score", "Ag_score_3", "Ag_score_6")])
  
plt_dat <- cd45neg_so %>%
  AddMetaData(clr_dat[c("Ag_score", "Ag_score_3", "Ag_score_6")])
  
# Format data for plots
ag_dat <- ag_dat %>%
  mutate_meta(
    mutate,
    Ag_score = case_when(
      exp == "exp-0" ~ ovalbumin_score,
      exp == "exp-1" & tm == 21 ~ ABPS4_score,
      exp == "exp-1" & tm == 42 ~ ABPS2_score,
      exp == "exp-2" & tm == 21 ~ ABPS2_score,
      exp == "exp-2" & tm == 42 ~ ABPS4_score
    ),
    Ag_score_3 = case_when(
      exp == "exp-1" & (tm == 21 | vaccination == "dual") ~ ABPS4_score,
      exp == "exp-2" & (tm == 21 | vaccination == "dual") ~ ABPS2_score
    ),
    Ag_score_6 = case_when(
      exp == "exp-1" & (tm == 42 | vaccination == "dual") ~ ABPS2_score,
      exp == "exp-2" & (tm == 42 | vaccination == "dual") ~ ABPS4_score
    ),
    mouse = fct_relevel(mouse, mice)
  )

plt_dat <- lec_so %>%
  AddMetaData(FetchData(ag_dat, c("Ag_score", "Ag_score_3", "Ag_score_6")))

plt_dat <- cd45neg_so %>%
  AddMetaData(FetchData(ag_dat, c("Ag_score", "Ag_score_3", "Ag_score_6")))

# Plot Ag-scores
plt_dat@meta.data %>%
  filter(vaccination == "single") %>%
  mutate(
    tm = fct_relevel(as.character(tm), tms),
    exp = ifelse(exp == "exp-0", "exp-1", exp)
  ) %>%
  ggplot(aes(tm, Ag_score, fill = subtype)) +
  geom_boxplot() +
  facet_grid(exp ~ subtype) +
  base_theme +
  theme(aspect.ratio = 1.5)

plt_dat@meta.data %>%
  # filter(tm == 21 | vaccination == "dual") %>%
  filter(tm == 42 | vaccination == "dual") %>%
  # ggplot(aes(as.character(tm), Ag_score_3, fill = vaccination)) +
  ggplot(aes(as.character(tm), Ag_score_6, fill = vaccination)) +
  geom_boxplot(notch = TRUE) +
  facet_grid(exp ~ subtype) +
  base_theme +
  theme(aspect.ratio = 1.5)
```

```{r "CELL TYPE FREQUENCY", eval = FALSE}
# Format data for plotting
typ_dat <- all_meta %>%
  filter(
    vaccination == "single",
    cell_type != "unassigned"
  ) %>%
  group_by(cell_sort) %>%
  mutate(
    tm        = fct_relevel(as.character(tm), tms),
    cell_type = fct_infreq(cell_type),
    cell_type = fct_relevel(cell_type, c("LEC", "Fibroblasts", "DC"))
  ) %>%
  ungroup() %>%
  arrange(tm, cell_type)

freq_dat <- typ_dat %>%
  mutate(
    tm_group = str_c(tm, "_", cell_sort),
    tm_group = fct_inorder(tm_group)
  )

# Create bargraphs
# these results are not very convincing since CD45 sorting varied significantly
# between our two experiments
group_lab <- c(CD45pos = "CD45+", CD45neg = "CD45-")

x_labs <- get_nlab_fun(freq_dat, "tm_group", l = "\n", r = "")
x_labs <- x_labs() %>%
  map_chr(~ str_remove_all(.x, "_.+"))

typ_freq <- freq_dat %>%
  ggplot(aes(tm_group, fill = cell_type, color = cell_type)) +
  geom_bar(position = "fill", key_glyph = draw_key_point) +
  facet_wrap(~ cell_sort, nrow = 1, scales = "free_x", labeller = as_labeller(group_lab)) +
  scale_fill_manual(values = typ_clrs) +
  scale_color_manual(values = typ_clrs) +
  scale_x_discrete(labels = x_labs) +
  scale_y_continuous(expand = expansion(c(0.01, 0.05))) +
  guides(fill = guide_legend(ncol = 2, override.aes = list(shape = 19, size = 4))) +
  labs(x = "days post vaccination") +
  base_theme +
  theme(
    aspect.ratio = 1.5,
    legend.title = element_blank(),
    panel.border = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x  = element_text(size = txt_pt2, angle = 45, hjust = 1, vjust = 1),
    axis.text.y  = element_blank(),
    axis.ticks   = element_blank()
  )
```

<br>

<br>

### Figure S1

A.  UMAP projections show cell types identified for CD45- and CD45+ samples.
B.  UMAP projections show LEC subsets identified for each timepoint.
C.  UMAP projections show DC subsets identified for each timepoint.
D.  The expression of key marker genes is shown for LEC subsets.
E.  The expression of key marker genes is shown for DC subsets.

```{r "type UMAPs"}
# Format data
typ_dat <- all_meta %>%
  mutate(
    vaccination = str_c(vaccination, "\nvaccination"),
    tm          = ifelse(is.na(tm), vaccination, str_c("day ", tm)),
    tm          = fct_relevel(tm, str_c("day ", tms)),
    cell_type   = fct_infreq(cell_type),
    cell_sort   = recode(cell_sort, CD45neg = "CD45-", CD45pos = "CD45+")
  )

n_dat <- typ_dat %>%
  group_by(tm, cell_sort) %>%
  summarize(n = str_c("n = ", label_comma()(n())), .groups = "drop")

# Create UMAP projections
typ_u <- typ_dat %>%
  plot_scatter(
    "cell_type",
    "hUMAP_1", "hUMAP_2",
    plot_colors  = typ_clrs,
    plot_lvls    = c("BEC", "LEC", "Epithelial cells", "Monocytes"),
    size         = 0.001,
    n_label      = "legend"
  ) +
  geom_text(
    aes(-Inf, Inf, label = n, color = NULL), data = n_dat,
    hjust = -0.1, vjust = 1.5, size = 10 / .pt
  ) +
  facet_grid(cell_sort ~ tm) +
  guides(color = guide_legend(nrow = 2, override.aes = list(size = 4))) +
  umap_theme_2 +
  theme(
    legend.position = "bottom",
    legend.title    = element_blank()
  )
```

```{r "subtype UMAPs"}
# Marker genes to plot
typ_gns <- list(
  LEC = c(
    "Lyve1", "Marco", "Madcam1",
    "Vcam1", "Ptx3",  "Ackr4"
  ),
  DC = c(
    "Xcr1", "Itgax", "Sirpa",
    "Lyz2", "Ccr7",  "Siglech"
  )
)

# Format data for subtype plots
subtype_dat <- ag_objs %>%
  imap(~ {
    clmns <- c("subtype", "tm", "vaccination", "hUMAP_1", "hUMAP_2")
    
    .x %>%
      FetchData(c(clmns, typ_gns[[.y]])) %>%
      as_tibble(rownames = ".cell_id") %>%
      mutate(
        vaccination = str_c(vaccination, "\nvaccination"),
        tm          = ifelse(is.na(tm), vaccination, str_c("day ", tm)),
        tm          = fct_relevel(tm, str_c("day ", tms)),
        subtype     = fct_relevel(subtype, levels(markers$annotation))
      )
  })

# Create UMAPs
subtype_u <- subtype_dat %>%
  imap(~ {
    .x %>%
      plot_scatter(
        "subtype", "hUMAP_1", "hUMAP_2",
        group_col    = "tm", panel_nrow = 2,
        n_label      = "corner",
        label_params = list(size = 14),
        size         = 0.1,
        plot_colors  = subtype_clrs[[.y]],
        plot_lvls    = levels(.x$subtype)
      ) +
      umap_theme_2 +
      theme(
        legend.position  = c(0.85, 0.25),
        legend.title     = element_blank()
      )
  })
```

```{r "subtype boxplots"}
# Create LEC/DC subtype boxplots
subtype_bxs <- subtype_dat %>%
  imap(~ {
    dat <- .x %>%
      pivot_longer(all_of(typ_gns[[.y]])) %>%
      mutate(
        subtype_tm = str_c(subtype, "_", tm),
        name       = fct_relevel(name, typ_gns[[.y]])
      ) %>%
      arrange(subtype) %>%
      mutate(subtype_tm = fct_inorder(subtype_tm))
    
    # Data for n labels
    n_labs <- dat %>%
      group_by(subtype, tm, subtype_tm) %>%
      summarize(
        n = label_comma()(n_distinct(.cell_id)),
        .groups = "drop"
      ) %>%
      mutate(
        n_lab = str_replace(subtype, "_", " "),
        n_lab = str_c(n_lab, " (", n, ")")
      )
    
    n_labs <- set_names(n_labs$n_lab, n_labs$subtype_tm)

    # Create boxplots
    n_typs <- n_distinct(dat$subtype)

    dat %>%
      ggplot(aes(subtype_tm, value, fill = subtype)) +
      geom_boxplot(outlier.size = 0.1, key_glyph = draw_key_point, alpha = 0.9) +
      facet_grid(name ~ tm, scales = "free") +
      guides(fill = guide_legend(override.aes = list(shape = 22, size = 4))) +
      scale_x_discrete(labels = n_labs) +
      scale_fill_manual(values = subtype_clrs[[.y]]) +
      base_theme +
      theme(
        aspect.ratio    = 4 / n_typs,
        legend.position = "none",
        strip.clip      = "off",
        axis.title      = element_blank(),
        axis.text.x     = element_text(angle = 90, hjust = 1, vjust = 0.5)
      )
  })
```

```{r "Fig S1", fig.width = 16, fig.height = 20}
figS1 <- typ_u /
  (subtype_u$LEC + subtype_u$DC) /
  wrap_plots(subtype_bxs$LEC, subtype_bxs$DC, widths = c(1, 0.63)) +
  plot_layout(heights = c(0.7, 0.7, 1)) +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(size = ttl_pt2 * 1.5, face = "bold"))

figS1
```

<br>

<br>

### Figure S2

A.  Ag-scores are shown for each timepoint for CD45- cell types. Two biological
    replicates are shown for the day 21 and day 42 timepoints. The number of
    cells identified for each cell type is shown above each timepoint.
B.  Ag-scores are shown for each timepoint for CD45+ cell types,
    as described in A.
C.  Ag-scores are shown for each timepoint for LEC subsets, as described in A.
D.  Ag-scores are shown for each timepoint for DC subsets, as described in A.

```{r "type Ag boxplots"}
# Format data for boxplots
dat <- c("exp-1", "exp-2") %>%
  map_dfr(~ {
    typ_ag_dat %>%
      filter(exp %in% c("exp-0", .x)) %>%
      mutate(
        exp       = str_replace(.x, "exp-", "r"),
        tm        = as.numeric(as.character(tm)),
        cell_sort = recode(cell_sort, CD45neg = "CD45-", CD45pos = "CD45+")
      )
  })

# Create boxplots
typ_ag_bxs <- c("CD45-", "CD45+") %>%
  set_names() %>%
  map(~ {
    dat <- dat %>%
      filter(cell_sort == .x)
      
    n_dat <- dat %>%
      group_by(cell_sort, exp, tm, cell_type) %>%
      summarize(n = label_comma()(n()), .groups = "drop")
    
    dat %>%
      create_ag_boxes(clrs = typ_clrs, facet_vars = c("exp", "cell_type")) +
      geom_text_repel(
        aes(y = Inf, label = n), data = n_dat,
        seed = 42, force = 0.005,
        size = 8 / .pt, vjust = 1.2
      ) +
      scale_y_continuous(expand = expansion(c(0.05, 0.15))) +
      theme(plot.margin = margin(15, 15, 30, 15))
  })
```

```{r "all subtype Ag boxplots"}
# Format data for boxplots
bx_dat <- ag_objs %>%
  map(~ {
    .x@meta.data %>%
      as_tibble(rownames = ".cell_id") %>%
      filter(vaccination == "single") %>%
      group_by(mouse, subtype, exp) %>%
      ungroup() %>%
      mutate(
        mouse   = as.character(mouse),
        tm_lab  = factor(as.character(tm), tms),
        subtype = fct_reorder(subtype, Ag_score, mean, .desc = TRUE)
      )
  })

# Create boxplots showing all LEC/DC subsets
all_ag_bxs <- bx_dat %>%
  imap(~ {
    dat <- .x
    
    dat <- c("exp-1", "exp-2") %>%
      map_dfr(~ {
        dat %>%
          filter(exp %in% c("exp-0", .x)) %>%
          mutate(exp = str_replace(.x, "exp-", "r"))
      })
    
    n_dat <- dat %>%
      group_by(tm, subtype, exp) %>%
      summarize(n = label_comma()(n()), .groups = "drop")
    
    dat %>%
      create_ag_boxes(clrs = subtype_clrs[[.y]], facet_vars = c("exp", "subtype")) +
      geom_text_repel(
        aes(y = Inf, label = n), data = n_dat,
        seed = 42, force = 0.005,
        size = 8 / .pt, vjust = 1.2
      ) +
      scale_y_continuous(expand = expansion(c(0.05, 0.15))) +
      theme(plot.margin = margin(15, 15, 30, 15))
  })
```

```{r "Fig S2", fig.width = 15, fig.height = 17.5}
# Create final figure
figS2 <- plot_grid(
  typ_ag_bxs$`CD45-`, typ_ag_bxs$`CD45+`,
  all_ag_bxs$LEC, all_ag_bxs$DC,
  labels = c("A", "B", "C", "D"),
  label_size = ttl_pt2 * 1.5,
  # rel_heights = c(1.1, 1.2, 1, 1.2),
  align = "v", axis = "rl",
  ncol = 1
)

figS2
```



```{r "OLD EXTENDED 1-2", eval = FALSE}
lec_dat %>%
  group_by(subtype, tm, exp) %>%
  summarize(n = n(), .groups = "drop")
```

```{r "OLD EXTENDED 1-3", fig.width = 9, fig.height = 6, eval = FALSE}
# Ag-score is shown for all DC subsets for each experiment

# Data for boxplots
dc_dat <- dc_so@meta.data %>%
  filter(mouse != "6wk-3wk") %>%
  mutate(subtype = fct_reorder(subtype, Ag_score, mean, .desc = TRUE))

ln_dat <- dc_dat %>%
  group_by(subtype, exp, tm) %>%
  summarize(Ag_score = median(Ag_score), .groups = "drop")

# Create boxplots
exps <- c("exp-1", "exp-2")

exps %>%
  map(~ {
    dc_dat %>%
      filter(exp %in% c("exp-0", .x)) %>%
      
      ggplot(aes(tm, Ag_score, fill = subtype)) +
      
      geom_line(
        data = filter(ln_dat, exp %in% c("exp-0", .x)),
        color = "grey85", linewidth = 0.5, linetype = 2
      ) +
      geom_boxplot(aes(group = mouse), outlier.size = 0.2, width = 3, key_glyph = draw_key_point) +
      
      guides(
        fill = guide_legend(override.aes = list(size = 4, shape = 22), order = 1),
        linetype = guide_legend(override.aes = list(color = "black"), title = NULL)
      ) +
      facet_grid(~ subtype) +
      
      scale_fill_manual(values = dc_clrs) +
      labs(title = .x, x = "days post vaccination", y = "Ag-score") +
      base_theme +
      theme(
        aspect.ratio    = 0.9,
        legend.position = "none",
        plot.margin     = margin(0, 15, 0, 15)
      )
  }) %>%
  plot_grid(
    plotlist = .,
    ncol = 1
  )

```

```{r "OLD EXTENDED 1-4", eval = FALSE}
dc_dat %>%
  group_by(subtype, tm, exp) %>%
  summarize(n = n(), .groups = "drop")
```

---

<br>

<br>

## Figure 2

Identification of an antigen archiving gene signature

A.  Ag-score is shown for day 14 Ag-high and -low cells identified for LEC
    subsets with the highest Ag-score.
B.  The fraction of cells predicted to be Ag-competent is shown for each LEC
    subset 14, 21, and 42 days post vaccination.
C.  Ag-high module score is shown for Ag-low, Ag-high, and predicted
    Ag-competent LECs.
    The Spearman correlation between Ag classes and Ag-high module score is
    shown for each timepoint.
    One-sided p values were calculated and adjusted using Bonferroni correction.
D.  UMAP projections show cLEC Ag-high module scores for each timepoint.
E.  The expression of select genes from the Ag-high (top four) and Ag-low
    (bottom four) gene modules is shown for cLECs. Genes identified as top
    predictors for all three LEC subsets are shown.

<u>Notes</u>

* Ag-low and -high cells were identified by separately clustering each d14 LEC
  subset for each sample into two groups based on Ag-score.
* The cutoffs identified for the d14 dataset were then used to identify Ag-high
  cells for the other timepoints.
* With this approach there is a different cutoff for each LEC subset, but the
  cutoff is consistent across timepoints.

```{r "RF Ag signal", fig.width = 12, fig.height = 3.25}
# Format plotting data
plt_dat <- lec_so@meta.data %>%
  filter(
    subtype %in% rf_typs,
    mouse %in% rf_mouse
  ) %>%
  mutate(
    mouse = as.character(mouse),  # to remove factor
    !!sym(rf_dat_clmn) := fct_relevel(!!sym(rf_dat_clmn), c("low", "high"))
  ) %>%
  arrange(!!sym(rf_dat_clmn))

# Format plot labels
plt_labs <- plt_dat %>%
  group_by(subtype, mouse, tm, !!sym(rf_dat_clmn)) %>%
  summarize(
    n = label_comma()(n()),
    .groups = "drop"
  ) %>%
  mutate(n = str_c(!!sym(rf_dat_clmn), ": ", n, " cells")) %>%
  group_by(subtype, mouse, tm) %>%
  arrange(!!sym(rf_dat_clmn)) %>%
  mutate(n = str_c(n, collapse = "\n"))

plt_clrs <- lec_clrs

# Create bargraphs
brs <- rf_typs %>%
  imap(~ {
    dat <- plt_dat %>%
      filter(subtype == .x) %>%
      mutate(!!sym(rf_dat_clmn) := fct_rev(!!sym(rf_dat_clmn)))
    
    dat %>%
      ggplot(aes(y = mouse, fill = subtype, alpha = !!sym(rf_dat_clmn))) +
      geom_bar() +
      facet_wrap(~ subtype, nrow = 1, scales = "free") +
      scale_alpha_manual(values = c(low = 0.35, high = 1)) +
      scale_fill_manual(values = plt_clrs) +
      scale_y_discrete(expand = expansion(c(0, 0))) +
      scale_x_continuous(expand = expansion(c(0, 0))) +
      theme_void() +
      theme(
        legend.position = "none",
        plot.margin     = margin(t = 5),
        strip.text      = element_text(size = ttl_pt2)
      )
  })

hsts <- rf_typs %>%
  imap(~ {
    dat <- plt_dat %>%
      filter(subtype == .x)
    
    fst <- .y == 1
    lst <- .y == length(rf_typs)
    
    # Create histograms
    hst <- dat %>%
      ggplot(aes(Ag_score, fill = subtype, alpha = !!sym(rf_dat_clmn))) +
      geom_histogram(bins = 30, key_glyph = draw_key_point) +
      geom_text(
        aes(-Inf, Inf, label = n),
        data = filter(plt_labs, subtype == .x),
        check_overlap = TRUE,
        alpha = 1,
        hjust = -0.25,
        vjust = 1.1,
        size  = txt_pt2 / .pt
      ) +
      facet_wrap(~ subtype, scales = "free", nrow = 1) +
      guides(fill = "none") +
      scale_alpha_manual(
        values = c(low = 0.35, high = 1),
        labels = function(x) str_c("Ag-", x)
      ) +
      scale_fill_manual(values = plt_clrs) +
      scale_y_continuous(expand = expansion(c(0.05, 0.25))) +
      guides(alpha = guide_legend(override.aes = list(shape = 22, size = 5))) +
      labs(x = "Ag-score", y = "number of cells") +
      base_theme +
      theme(
        aspect.ratio    = 0.6,
        strip.text      = element_blank(),
        legend.position = "none",
        legend.title    = element_blank(),
        plot.margin     = margin(t = 0, r = 5, b = 0, l = 5)
      )
    
    if (!fst && !lst) {
      hst <- hst +
        theme(legend.position = "bottom")
    }
    
    if (!fst) {
      hst <- hst +
        theme(axis.title.y = element_blank())
    }
    
    if (fst || lst) {
      hst <- hst +
        theme(axis.title.x = element_blank())
    }
    
    hst
  })

# Create final figure
hsts <- wrap_plots(
  append(brs, hsts),
  ncol = length(rf_tms),
  heights = c(0.1, 1)
)
```

```{r "RF Ag-competent bars", fig.width = 8, fig.height = 5}
# Format data for bargraphs
rf_bar_dat <- lec_so@meta.data %>%
  filter(
    tm %in% rf_tms,
    subtype %in% rf_typs,
    !training_data
  ) %>%
  mutate(
    subtype    = fct_relevel(subtype, rf_typs),
    rf_correct = fct_relevel(as.character(rf_correct), c("TRUE", "FALSE")),
    pred_grp   = fct_relevel(pred_grp, rev(names(pred_clrs))),
    tm         = as.character(tm)
  )

bar_thm <- base_theme +
  theme(aspect.ratio = 1.4)

rf_bar_1 <- rf_bar_dat %>%
  ggplot(aes(tm, fill = pred_grp)) +
  geom_bar(position = "fill", alpha = 0.85, key_glyph = draw_key_point) +
  facet_grid(~ subtype) +
  scale_fill_manual(values = pred_clrs) +
  scale_y_continuous(breaks = c(0, 0.5, 1)) +
  guides(fill = guide_legend(override.aes = list(shape = 22, size = 5))) +
  labs(y = "fraction of cells", x = "days post vaccination") +
  bar_thm +
  theme(
    legend.title = element_blank(),
    legend.position = "bottom"
  )
```

```{r "RF module boxplots", fig.width = 10, fig.height = 4}
# Format plot data
# exclude dual vaccination data
clmns <- c(
  "tm", "subtype", "pred_grp", "training_data",
  "Ag_score", ag_modules
)

rf_bx_dat <- lec_so %>%
  FetchData(clmns) %>%
  as_tibble(rownames = ".cell_id") %>%
  filter(
    tm %in% rf_tms
    # !training_data
  )

rf_mod_args <- list(
  mod_typ = c("high", "low"),
  p_alt   = c("greater", "less"),
  p_hjust = c(0.7, 0.55)
)

mod_bxs <- rf_mod_args %>%
  pmap(~ {
    args <- list(...)
    
    module_clmns <- str_c(rf_typs, "_", args$mod_typ)
    module_title <- str_c("Ag-", args$mod_typ, " module score")
    
    rf_bx_dat %>%
      create_rf_module_boxes(
        module_clmns = module_clmns,
        module_title = module_title,
        p_alt        = args$p_alt,
        p_hjust      = c(args$p_hjust, 0.5),
        p_vjust      = c(1.2, 2),
        clrs         = pred_clrs
      )
  })
```

```{r "RF cLEC module UMAPs", fig.width = 8, fig.height = 6.5}
# Format plotting data
u_dat <- lec_so@meta.data %>%
  as_tibble(rownames = ".cell_id") %>%
  filter(tm %in% rf_tms) %>%
  mutate(Ag_class_ml = str_c("Ag-", Ag_class_ml))

# Create UMAPs
u_args <- list(
  typ = set_names(rf_typs),
  pt_size = list(1, c(0.5, 0.1), c(0.5, 0.1))
)

mod_u <- u_args %>%
  pmap(~ {
    u_dat %>%
      create_rf_module_umaps(
        typ           = .x,
        ag_class_clmn = "Ag_class_ml",
        module_clmn   = str_c(.x, "_high"),
        module_title  = str_c(.x, " Ag-high module"),
        equal_scales  = FALSE,
        pt_size       = .y
      )
  })
```

```{r "RF top features examples", fig.height = 8, fig.width = 8}
# # Calculate p-values WIP
# rf_feats %>%
#   imap_dfr(~ {
#     typ <- .y
#     
#     .x %>%
#       map_dfr(~ {
#         dat <- lec_so %>%
#           subset(subtype == typ) %>%
#           subset(tm %in% rf_tms) %>%
#           # subset(training_data == FALSE) %>%  # WHY CAN'T SUBSET HANDLE `!` ??
#           FetchData(c("pred_grp", "tm", .x)) %>%
#           as_tibble(rownames = "cell_id") %>%
#           pivot_longer(all_of(.x))
#       })
#   })

# Identify genes shared between multiple cell types
typ <- "cLEC"

shared_feats <- c("high", "low") %>%
  set_names() %>%
  map(~ {
    ag <- .x
    
    fts <- rf_feats[rf_typs] %>%
      map(pluck, ag)
    
    fts <- fts %>%
      map(~ {
        .x[.x %in% rf_feats$cLEC[[ag]]]
      })
    
    fts <- fts[!names(fts) %in% c(typ, "BEC")]
    
    fts %>%
      purrr::reduce(c) %>%
      unique() %>%
      sort()
  })

# Plot example genes
gene_examples <- lec_so %>%
  subset(subtype == rf_cell_type) %>%
  subset(tm %in% rf_tms) %>%
  create_gn_plots(
    ag_high_gns = shared_feats$high,
    ag_low_gns  = shared_feats$low,
    plt_clrs    = pred_clrs,
    x_lvls      = names(pred_clrs),
    n_gns       = 4,
    
    # top_gns   = c("Gdf10", "Tgfa", "Ctsd", "Grn", "Col4a1", "Col4a2", "Fau", "Lama4"),
    
    p_test    = wilcox.test,
    draw_line = FALSE,
    pt_size   = 2.5
  ) +
  ggtitle(str_c(rf_cell_type, " expression")) +
  theme(plot.title = element_text(size = 18, hjust = 0.5))
```

```{r "Fig 2", fig.width = 15, fig.height = 16.1}
left <- plot_grid(
  plot_grid(hsts, nrow = 2, rel_heights = c(1, 0.1)),
  mod_bxs[[1]],
  mod_u$cLEC[[1]],
  mod_u$cLEC[[2]],
  ncol = 1,
  rel_heights = c(1, 1, 1.11, 1.1),
  labels = c("A", "C", "D"),
  label_size = ttl_pt2 * 1.5,
  label_y = c(1, 1.1, 1)
)

right <- plot_grid(
  rf_bar_1, gene_examples,
  ncol = 1,
  rel_heights = c(1, 3),
  labels = c("B", "E"),
  label_size = ttl_pt2 * 1.5
)

fig2 <- plot_grid(
  left, right,
  rel_widths = c(1, 0.5)
)

fig2
```



```{r "RF GO", eval = FALSE}
# # Load GO results
# go <- qread(here(params$mod_dir, "ag-high_go.qs"))
# 
# endo_gns <- go@compareClusterResult %>%
#   filter(
#     Description == "endocytosis",
#     p.adjust < 0.05
#   ) %>%
#   pull(geneID) %>%
#   str_split("/") %>%
#   unlist()
# 
# gns <- c("Ctsd", "Dab2", "Prcp", "Cltc", "Degs", "Clu", "Atp6v1a", endo_gns)
# 
# gns <- unique(gns[gns %in% rf_feats$cLEC$high])
# 
# lec_so %>%
#   subset(subtype == "cLEC") %>%
#   subset(mouse == "3wk") %>%
#   mutate_meta(
#     mutate,
#     Ag_class_ml = ifelse(is.na(pred_grp), pred_grp, Ag_class_ml)
#   ) %>%
#   plot_violin(
#     "Atp6v1a",
#     "pred_grp",
#     # "Ag_class_ml",
#     "mouse",
#     method = "boxplot"
#   )
# 
# dat <- lec_so %>%
#   FetchData(c("Atp6v1a", "pred_grp", "subtype", "mouse")) %>%
#   filter(mouse == "3wk", subtype == "cLEC")
# 
# wilcox.test(
#   dat %>% filter(pred_grp == "Ag-high") %>% pull(Atp6v1a),
#   dat %>% filter(pred_grp == "Ag-competent") %>% pull(Atp6v1a)
# )
```

<br>

<br>

### Figure S3

A.  The fraction of cells predicted to be Ag-low and Ag-high is shown for each
    LEC subset.
C.  Ag-low module score is shown for Ag-low, Ag-high, and predicted
    Ag-competent LECs.
    The Spearman correlation between Ag classes and Ag-high module score is
    shown for each timepoint.
    One-sided p values were calculated and adjusted using Bonferroni correction.
D.  UMAP projections show fLEC Ag-high module scores for each timepoint.
E.  UMAP projections show collecting LEC Ag-high module scores for each
    timepoint.

```{r "RF predicted Ag class", fig.width = 8, fig.height = 5}
# Plot fraction predicted groups
rf_bar_2 <- rf_bar_dat %>%
  ggplot(aes(tm, fill = rf_pred, alpha = rf_correct)) +
  geom_bar(position = "fill") +
  scale_alpha_manual(values = c(`TRUE` = 1, `FALSE` = 0.5)) +
  scale_fill_manual(values = pred_clrs) +
  scale_y_continuous(breaks = c(0, 0.5, 1)) +
  facet_grid(~ subtype) +
  labs(y = "fraction of cells", x = "days post vaccination") +
  guides(
    fill = guide_legend(order = 1, title = "prediction"),
    alpha = guide_legend(title.theme = element_blank())
  ) +
  bar_thm
```

```{r "Fig S3", fig.width = 20, fig.height = 12.5}
top <- plot_grid(
  rf_bar_2, mod_bxs[[2]],
  rel_widths = c(0.75, 1),
  labels = c("A", "B"),
  label_size = ttl_pt2 * 1.5,
  align = "h",
  axis  = "tb"
)

left <- plot_grid(
  plotlist = mod_u$fLEC,
  ncol  = 1,
  align = "v",
  axis  = "rl"
)

right <- plot_grid(
  plotlist = mod_u$Collecting,
  ncol  = 1,
  align = "v",
  axis  = "rl"
)

bot <- plot_grid(
  left, right,
  labels = c("C", "D"),
  label_size = ttl_pt2 * 1.5,
  nrow  = 1,
  align = "h",
  axis  = "tb"
)

figS3 <- plot_grid(
  top, bot,
  ncol = 1,
  rel_heights = c(0.5, 1)
)

figS3
```

---

<br>

<br>

## Figure 3

Antigen archiving is enhanced by successive vaccinations

A.  Mean 21 day Ag-score is shown for LECs from mice that received a single
    vaccination (21 day or 42 day) or dual vaccination (21 day and 42 day).
B.  UMAP projections show 21 day Ag-scores for LEC subsets for single and dual
    vaccinations.
C.  Prior vaccination enhances antigen archiving.
    Ag-score is shown for single and dual vaccinations for the 21 day timepoint
    for each LEC subset. Other timepoints are shown in grey. P values
    were calculated using a one-sided Wilcoxon rank sum test with Bonferroni
    correction.
D.  Mean 42 day Ag-score is shown as described in A.
E.  UMAP projections show 42 day Ag-scores for LEC subsets as described in B.
F.  Successive vaccinations enhances retention of previously archived antigen.
    Ag-score is shown for the 42 day timepoint as described in C.
G.  21 day and 42 day Ag-score is compared for LEC subsets.
H.  Ag-high module scores described in Figure 2 are shown for LECs that archived
    antigens from both vaccinations (double-high), from only one vaccination
    (single-high), or have low levels of both antigens (Ag-low). Module scores
    are shown for the corresponding LEC subset. P values were calculated using
    a one-sided Wilcoxon rank sum test with Bonferroni correction.
I.  Ag-low module scores are shown for LEC subset as described in H.

```{r "exchange UMAPs", fig.width = 16, fig.height = 4}
# Format Ag exchange plot data
# use experiment 1 since, experiment 2 is missing 3wk LECs
ex_dat <- lec_so@meta.data %>%
  as_tibble(rownames = ".cell_id") %>%
  filter(exp %in% c("exp-0", ag_exps$LEC)) %>%
  mutate(
    mouse = m_key[as.character(mouse)],
    mouse = fct_relevel(mouse, unname(m_key))
  )

# Create UMAPs
ex_heat_args <- list(
  ag_clmns = ag_clmns,
  clr      = tm_clrs[c("21", "42")]
)

ex_u <- ex_heat_args %>%
  pmap(~ {
    dat <- ex_dat %>%
      filter(exp == ag_exps$LEC)
    
    # mx_sig <- max(dat[ex_heat_args$ag_clmns])
    
    dat %>%
      plot_scatter(
        ..1,
        x           = "hUMAP_1",
        y           = "hUMAP_2",
        group_col   = "mouse",
        panel_nrow  = 1,
        outline     = TRUE,
        size        = 0.6,
        stroke      = 0.7,
        plot_colors = c("lightblue", "white", ..2),
        label_params = list(size = ttl_pt1)
      ) +
      # scale_fill_gradientn(colours = c("lightblue", "white", ..2), limits = c(0, mx_sig)) +
      guides(fill = guide_colorbar(
        ticks = FALSE,
        title = ag_labs[.x],
        barheight = unit(150, "pt"), barwidth = unit(10, "pt")
      )) +
      umap_theme +
      theme(
        legend.position = "right",
        panel.border    = element_rect(color = ln_col, linewidth = ln_pt)
      )
  })
```

```{r "exchange heatmaps"}
# Format data for heatmaps
heat_dat <- ex_dat %>%
  filter(
    exp == ag_exps$LEC,
    subtype %in% ag_plt_typs$LEC
  ) %>%
  group_by(mouse, subtype) %>%
  summarize(across(all_of(ag_clmns), mean), .groups = "drop")

# Create heatmaps
ex_heat <- ex_heat_args %>%
  pmap(~ {
    heat_dat %>%
      create_exchange_heats(
        ag_clmn = .x,
        ttl     = ag_labs_2[[.x]],
        clr     = .y
      )
  })
```

```{r "exchange boxes"}
# Effect of successive vaccinations on antigen uptake/retention
# use LEC subsets shown in Fig 1
ex_bx_args <- list(
  tm      = c("21", "42"),
  tm_clmn = c("tm_3", "tm_6"),
  ag_clmn = c("Ag_score_3", "Ag_score_6"),
  p_hjust = c(0.5, 0.8)
)

ex_bxs <- ex_bx_args %>%
  pmap(~ {
    args <- list(...)
    
    ex_dat %>%
      create_exchange_boxes(
        tm        = args$tm,
        tm_clmn   = args$tm_clmn,
        ag_clmn   = args$ag_clmn,
        typs      = ag_plt_typs$LEC,
        clrs      = tm_clrs,
        p_cutoff = Inf,
        p_hjust   = args$p_hjust,
        p_vjust   = 0.7
      )
  })
```

```{r "exchange correlation", fig.width = 10, fig.height = 4}
# Format scatter plot data
scat_dat <- lec_so@meta.data %>%
  as_tibble(rownames = ".cell_id") %>%
  filter(
    vaccination == "dual",
    subtype %in% ag_plt_typs$LEC,
  ) %>%
  mutate(subtype = fct_relevel(subtype, ag_plt_typs$LEC))

ex_scat <- scat_dat %>%
  split(.$exp) %>%
  map(create_exchange_scatter)
```

```{r "exchange Ag-high module scores", fig.width = 9, fig.height = 5.5}
# Format plot data
# module scores for biological replicates are similar
# pool all cells together to increase power
clmns <- c(
  "exp", "subtype", "mouse", "vaccination",
  "rf_pred", "Ag_class_dual", ag_modules
)

ex_mod_bx_dat <- lec_so %>%
  FetchData(clmns) %>%
  filter(vaccination == "dual") %>%
  mutate(Ag_class_dual = recode(Ag_class_dual, low = "Ag-low")) %>%
  
  mutate(Ag_class_dual = ifelse(
    Ag_class_dual == "Ag-low" & rf_pred == "Ag-high",
    "Ag-competent",
    Ag_class_dual)
  )

# Create boxplots
bx_lvls <- c("Ag-low", "Ag-competent", "single-high", "double-high")

rf_mod_args$p_x <- c(1, 4)
rf_mod_args$p_hjust <- c(0.2, 0.8)

ex_mod_bxs <- rf_mod_args %>%
  pmap(~ {
    args <- list(...)

    module_clmns <- str_c(rf_typs, "_", args$mod_typ)
    module_title <- str_c("Ag-", args$mod_typ, " module score")

    ex_mod_bx_dat %>%
      create_module_boxes(
        module_clmns = module_clmns,
        module_title = module_title,
        lvls         = bx_lvls,
        p_alt        = args$p_alt,
        p_x          = args$p_x,
        p_hjust      = args$p_hjust
      ) +
      theme(plot.margin = margin(15, 15, 5, 15))
  })
```

```{r "Fig 3", fig.width = 15.5, fig.height = 19}
fig3 <- (ex_heat[[1]] + ex_u[[1]]) / ex_bxs[[1]] /
  (ex_heat[[2]] + ex_u[[2]]) / ex_bxs[[2]] /
  ex_scat[[ag_exps$LEC]] /
  (ex_mod_bxs[[1]] + ex_mod_bxs[[2]]) +
  plot_layout(heights = c(1, 0.55, 1, 0.55, 0.8, 0.95)) +
  plot_annotation(tag_levels = "A") &
  theme(
    plot.tag = element_text(size = ttl_pt2 * 1.5, face = "bold"),
    plot.tag.position = c(0, 1)
  )

fig3
```

<br>

<br>

### Figure S4

A.  Ag-score is shown for single and dual vaccinations for the 21 day timepoint
    for DC subsets from a biological replicate.
    This analysis was not performed for a biological replicate for LEC subsets
    due to the low number of LECs recovered (<40 total cells).
    Other timepoints are shown in grey.
    P values were calculated using a one-sided t test with Bonferroni
    correction.
B.  Ag-score is shown for the 42 day timepoint for LEC subsets from a
    biological replicate, as described in A.
C.  Ag-score is shown for the 42 day timepoint for DC subsets from a biological
    replicate, as described in A.
D.  21 day and 42 day Ag-score is compared for LEC subsets from a biological
    replicate.

```{r "rep exchange boxes"}
# Set input data
# plot replicates not shown in main figure
exps        <- c("exp-1", "exp-2")
plt_lec_exp <- exps[exps != ag_exps$LEC]
plt_dc_exp  <- exps[exps != ag_exps$DC]

ex_bx_args$exps <-list(
  list(LEC = NULL,        DC = plt_dc_exp),
  list(LEC = plt_lec_exp, DC = plt_dc_exp)
)

# Create boxplots for biological replicates
rep_ex_bxs <- ex_bx_args %>%
  pmap(~ {
    args <- list(...)
    
    objs <- purrr::discard(args$exps, is.null)
    objs <- ag_objs[names(objs)]
    
    objs %>%
      imap(~ {
        dat <- .x@meta.data
        
        dat <- args$exps[[.y]] %>%
          map_dfr(~ {
            dat %>%
              filter(exp %in% c("exp-0", .x))
          })
        
        dat %>%
          create_exchange_boxes(
            tm         = args$tm,
            tm_clmn    = args$tm_clmn,
            ag_clmn    = args$ag_clmn,
            clrs       = tm_clrs,
            typs       = ag_plt_typs[[.y]],
            p_hjust    = args$p_hjust,
            facet_vars = "subtype",
            p_cutoff   = Inf
          ) +
          scale_y_continuous(expand = expansion(c(0.05, 0.15))) +
          theme(legend.position = "bottom")
      })
  }) %>%
  flatten()
```

```{r "RF module exchange correlation", fig.height = 8}
# Set plot colors
ex_ag_clrs <- c(
  `double-high` = "#D7301F",
  `single-high` = "black",
  `Ag-low`      = "#56B4E9"
  # `single-high` = "#6A51A3",
)

# Format plotting data
scat_dat <- lec_so@meta.data %>%
  as_tibble(rownames = ".cell_id") %>%
  filter(vaccination == "dual") %>%
  mutate(Ag_class_dual = recode(Ag_class_dual, low = "Ag-low"))

# Create scatter plots
scat_args <- list(
  mod_typ     = c("high", "low"),
  label_x     = c(-Inf, Inf),
  label_hjust = c(-0.1, 1.1)
)

ex_mod_scat <- scat_args %>%
  pmap(~ {
    args <- list(...)
    
    module_clmns <- str_c(rf_typs, "_", args$mod_typ)
    module_title <- str_c("Ag-", args$mod_typ, " module score")
    
    scat_dat %>%
      create_module_scatter(
        ag_clmns     = ag_clmns,
        module_clmns = module_clmns,
        clrs         = ex_ag_clrs,
        module_title = module_title,
        label_x      = args$label_x,
        label_hjust  = args$label_hjust
      )
  })
```

```{r "Fig S4", fig.width = 13, fig.height = 15}
# Create final figure
figS4 <- wrap_plots(rep_ex_bxs, ncol = 1) /
  ex_scat$`exp-2` /
  (ex_mod_scat[[1]] + ex_mod_scat[[2]]) +
  plot_layout(heights = c(1, 1, 1, 1.5, 3)) +
  plot_annotation(tag_levels = "A") &
  theme(
    plot.tag = element_text(size = ttl_pt2 * 1.5, face = "bold"),
    plot.tag.position = c(0, 1)
  )

figS4
```



```{r "EXCHANGE SCATTER PLOTS EXTENDED", fig.width = 14, fig.height = 4, eval = FALSE}
# Plot data
dat <- lec_so@meta.data %>%
  as_tibble(rownames = "cell_id") %>%
  filter(
    exp == "exp-1",
    mouse == "6wk-3wk",
  ) %>%
  # group_by(subtype) %>%
  # filter(n() > 50) %>%
  # ungroup() %>%
  mutate(subtype = fct_relevel(subtype, ag_plt_typs$LEC))

# Create scatter plots
x_var <- "Ag_6wk_score"
y_var <- "Ag_3wk_score"

ag_labs <- c(
  Ag_3wk_score = "Ag-score (21 day)",
  Ag_6wk_score = "Ag-score (42 day)"
)

dat %>%
  ggplot(aes(!!sym(x_var), !!sym(y_var), color = Ag_class_dual)) +
  geom_abline(slope = 1, intercept = 0, linewidth = 0.5, color = "grey90") +
  geom_point(size = 1.5) +
  geom_smooth(
    method    = "lm",
    linetype  = 2,
    linewidth = 0.5,
    color     = "black",
    se        = FALSE,
    formula   = "y ~ x"
  ) +
  facet_wrap(~ subtype, nrow = 1) +
  coord_cartesian(
    xlim = c(0, max(c(dat[[x_var]], dat[[y_var]]))),
    ylim = c(0, max(c(dat[[x_var]], dat[[y_var]])))
  ) +
  labs(
    x = ag_labs[x_var],
    y = ag_labs[y_var]
  ) +
  base_theme +
  theme(
    aspect.ratio = 1,
    legend.position = "bottom",
    legend.title    = element_blank()
  )
```

```{r "AG-EX DOUBLE POSITIVE EXTENDED", fig.width = 14, fig.height = 3, eval = FALSE}
# Expression of Ag-high gene modules is shown for each LEC subset. Expression is
# similar for the day 42 cLEC Ag-low/high groups, likely due to the many Ag-low
# cells that are actually Ag-competent and should express the Ag-high gene
# module.
# 
# Format plot data
lvls <- c("Ag-low", "Ag-high", "single-high", "double-high")

plt_dat <- lec_so %>%
  FetchData(c("subtype", "mouse", "Ag_class_dual", ag_modules)) %>%
  filter(
    mouse %in% c("6wk", "3wk", "6wk-3wk"),
    subtype %in% rf_typs
  ) %>%
  pivot_longer(all_of(ag_modules)) %>%
  separate_wider_regex(
    name,
    c(module_type = ".*", "_", module_ag = ".*")
  ) %>%
  filter(
    subtype == module_type,
    module_ag == "high"
  ) %>%
  
  # Group all low cells together under single mouse, otherwise three boxplots
  # are shown
  mutate(
    subtype       = fct_relevel(subtype, rf_typs),
    mouse         = m_key[as.character(mouse)],
    Ag_class_dual = str_replace(Ag_class_dual, "^[0-9]+wk", "Ag"),
    Ag_class_dual = recode(Ag_class_dual, low = "Ag-low"),
    Ag_class_dual = fct_relevel(Ag_class_dual, lvls)
  ) %>%
  arrange(Ag_class_dual) %>%
  mutate(
    mouse_class = fct_inorder(str_c(mouse, "-", subtype, "-", Ag_class_dual))
  )

# Format n labels
n_dat <- plt_dat %>%
  group_by(Ag_class_dual, mouse_class, subtype, module_ag) %>%
  summarize(n = n(), .groups = "drop") %>%
  mutate(
    n = str_c(Ag_class_dual, "\nn = ", scales::label_comma()(n))
  ) %>%
  filter(module_ag == "high")

n_lab <- set_names(n_dat$n, n_dat$mouse_class)

# Create boxplots
ex_mod_bxs <- rf_typs %>%
  map(~ {
    plt_dat %>%
      filter(subtype == .x) %>%
      ggplot(aes(mouse_class, value, fill = subtype)) +
      geom_boxplot(notch = TRUE, outlier.size = 0.1) +
      facet_grid(~ mouse, scales = "free_x", space = "free_x") +
      scale_fill_manual(values = lec_clrs) +
      scale_x_discrete(labels = n_lab) +
      labs(y = str_c(.x, " Ag-high\nmodule score")) +
      base_theme +
      theme(
        legend.position = "none",
        axis.text.x  = element_text(angle = 90, hjust = 1, vjust = 0.5),
        axis.title.x = element_blank()
      )
  })

ex_mod_bxs %>%
  plot_grid(
    plotlist = .,
    align    = "h",
    axis     = "tb",
    nrow     = 1
  )
```

---

<br>

<br>

## Figure 4

Antigen uptake by DCs is enhanced by successive vaccinations

A.  Mean 21 day Ag-score is shown for DCs from mice that received a single
    vaccination (21 day or 42 day) or dual vaccination (21 day and 42 day).
B.  Ag-score is shown for single and dual vaccinations for the 21 day timepoint
    for DC subsets with highest Ag-score. Other timepoints are shown in grey.
    P values were calculated using a one-sided Wilcoxon rank sum test with
    Bonferroni correction.
C.  Mean 42 day Ag-score is shown for DCs from mice that received single or
    dual vaccination as described in A.
D.  Ag-score is shown for single and dual vaccinations for the 42 day timepoint
    as described in B.
    
<u>Notes</u>

* Need to add GeoMx figures showing Ag signal is highest in medula and sinus
  regions for both LECs and DCs.

```{r "DC exchange heatmaps", fig.width = 16, fig.height = 4}
# Format data for DC exchange plots
dc_ex_dat <- dc_so@meta.data %>%
  as_tibble(rownames = ".cell_id") %>%
  filter(
    exp %in% c("exp-0", ag_exps$DC),
    subtype %in% ag_plt_typs$DC
  ) %>%
  mutate(
    mouse = m_key[as.character(mouse)],
    mouse = fct_relevel(mouse, unname(m_key))
  )

# Format data for heatmaps
dc_heat_dat <- dc_ex_dat %>%
  filter(exp == ag_exps$DC) %>%
  group_by(mouse, subtype) %>%
  summarize(across(all_of(ag_clmns), mean), .groups = "drop")

# Create heatmaps
dc_ex_heat <- ex_heat_args %>%
  pmap(~ {
    dc_heat_dat %>%
      create_exchange_heats(
        ag_clmn = .x,
        ttl     = ag_labs_2[[.x]],
        clr     = .y
      )
  })
```

```{r "DC exchange boxes", fig.width = 10, fig.height = 2}
# Effect of successive vaccinations on antigen uptake/retention
# use DC subsets shown in figure 1
dc_ex_bxs <- ex_bx_args %>%
  pmap(~ {
    args <- list(...)
    
    dc_ex_dat %>%
      create_exchange_boxes(
        tm        = args$tm,
        tm_clmn   = args$tm_clmn,
        ag_clmn   = args$ag_clmn,
        typs      = ag_plt_typs$DC[1:3],
        clrs      = tm_clrs,
        p_size    = 11,
        p_hjust   = args$p_hjust,
        p_vjust   = 0.7
      ) +
      theme(legend.position = "bottom")
  })
```

```{r "Fig 4", fig.width = 14, fig.height = 7}
fig4 <- wrap_plots(
  dc_ex_heat[[1]], dc_ex_bxs[[1]],
  dc_ex_heat[[2]], dc_ex_bxs[[2]],
  nrow = 2,
  widths = c(0.25, 1)
) +
  plot_annotation(tag_levels = "A") &
  theme(
    plot.tag = element_text(size = ttl_pt2 * 1.5, face = "bold"),
    plot.tag.position = c(0, 1)
  )

fig4
```



```{r "EXTENDED 4-1", eval = FALSE}

dat %>%
  plot_scatter(
    # "Ag_6wk_score",
    "Ag_3wk_score",
    x           = "hUMAP_1",
    y           = "hUMAP_2",
    group_col   = "mouse",
    panel_nrow  = 1,
    # outline     = TRUE,
    size        = 0.5,
    stroke      = 0.7,
    # plot_colors = c("lightblue", "white", "#D7301F"),
    plot_colors = c("lightblue", "white", rep("#6A51A3", 3)),
    # max_q       = 99.9,
    label_params = list(size = ttl_pt1)
  ) +
  # scale_fill_gradientn(colours = c("lightblue", "white", ..2), limits = c(0, mx_sig)) +
  guides(fill = guide_colorbar(
    ticks = FALSE,
    # title = ag_labs_2[.x],
    barheight = unit(150, "pt"), barwidth = unit(10, "pt")
  )) +
  umap_theme +
  theme(
    legend.position = "right",
    panel.border    = element_rect(color = ln_col, linewidth = ln_pt)
  )

```

```{r "EXTENDED 4-2", eval = FALSE}

# Create UMAPs
plt_args <- list(
  ag_clmns = ag_clmns,
  clr      = tm_clrs[c("21", "42")]
)

mx_sig <- max(dat[plt_args$ag_clmns])

ex_u <- plt_args %>%
  pmap(~ {
    dat %>%
      plot_scatter(
        ..1,
        x           = "hUMAP_1",
        y           = "hUMAP_2",
        group_col   = "mouse",
        panel_nrow  = 1,
        outline     = TRUE,
        size        = 1,
        stroke      = 0.7,
        plot_colors = c("lightblue", "white", ..2),
        max_q       = 99,
        label_params = list(size = ttl_pt1)
      ) +
      # scale_fill_gradientn(colours = c("lightblue", "white", ..2), limits = c(0, mx_sig)) +
      guides(fill = guide_colorbar(
        ticks = FALSE,
        title = ag_labs_2[.x],
        barheight = unit(150, "pt"), barwidth = unit(10, "pt")
      )) +
      umap_theme +
      theme(
        legend.position = "right",
        panel.border    = element_rect(color = ln_col, linewidth = ln_pt)
      )
  })

```

---

<br>

<br>

## Figure 5

Antigen archiving is impaired during CHIKV infection

A.  UMAP projections show LEC subsets for mock and CHIKV-infected mice.
B.  UMAP projections show predicted Ag-competent cLECs for mock and
    CHIKV-infected mice.
C.  The fraction of predicted Ag-competent cLECs is shown for mock and
    CHIKV-infected mice for each biological replicate.
D.  UMAP projections show cLEC Ag-high module scores for mock and CHIKV-infected
    mice.
E.  cLEC Ag-high module scores are shown for mock and CHIKV-infected mice for
    each biological replicate.
F.  Expression of the cLEC Ag-high gene module is shown for mock and CHIKV-infected
    mice for cLECs from each biological replicate.
    
<u>Notes</u>

* Show cLEC Ag-low results in main figure?

```{r "CHIKV data"}
# CHIKV theme
chikv_clrs <- c(
  `Ag-low`       = "#56B4E9",
  `Ag-competent` = "#D7301F",
  other          = "white"
)

treat_clrs <- set_names(chikv_clrs[1:2], c("mock", "CHIKV"))

chikv_u_theme <- umap_theme_2 +
  theme(aspect.ratio = 0.7)

# Load CHIKV data
rm(lec_so, dc_so)

chikv_so <- qread(here(params$chikv_dir, "so_lec.qs"))

chikv_so <- chikv_so %>%
  subset(tm == "24hpi") %>%
  AddMetaData(chikv_preds, col.name = str_c("rf_", colnames(chikv_preds)))

chikv_so <- chikv_so %>%
  mutate_meta(
    mutate,
    pred_grp  = dplyr::recode(rf_pred, high = "Ag-competent", low = "Ag-low"),
    pred_grp  = fct_relevel(pred_grp, rev(names(chikv_clrs))),
    treatment = fct_relevel(treatment, c("mock", "CHIKV")),
    subtype   = lec_subtype  # add subtype column to keep consistent with other
  )                          # objects

chikv_so <- chikv_so %>%
  AddMetaData(FetchData(., c("hUMAP_1", "hUMAP_2")))

# Add module scores
rf_feats %>%
  iwalk(~ {
    clmn_nms <- str_c(.y, "_", names(.x))

    clmn_nms <- set_names(
      str_c(clmn_nms, seq_along(clmn_nms)),
      clmn_nms
    )

    chikv_so <<- chikv_so %>%
      AddModuleScore(
        features = .x,
        name     = names(clmn_nms)
      ) %>%
      mutate_meta(dplyr::rename, !!!syms(clmn_nms))
  })

# Data for plotting
chikv_dat <- chikv_so@meta.data %>%
  as_tibble(rownames = ".cell_id")
```

```{r "CHIKV subsets"}
# LEC subsets
top <- chikv_so %>%
  plot_scatter(
    "lec_subtype",
    x            = "hUMAP_1",
    y            = "hUMAP_2",
    group_col    = "treatment",
    plot_colors  = lec_clrs,
    size         = 0.75,
    label_params = list(size = ttl_pt2)
  ) +
  chikv_u_theme +
  theme(
    legend.title = element_blank(),
    panel.border = element_blank()
  )
```

```{r "CHIKV Ag class UMAPs"}
# Create UMAPs
u_args <- list(
  typ = set_names(rf_typs),
  pt_size = list(1, c(0.5, 0.1), c(0.5, 0.1))
)

chikv_u <- u_args %>%
  pmap(~ {
    chikv_dat %>%
      create_rf_module_umaps(
        typ           = .x,
        ag_class_clmn = "pred_grp",
        module_clmn   = str_c(.x, "_high"),
        module_title  = str_c(.x, "\nAg-high module"),
        facet_var     = "treatment",
        lab_fn        = "label_value",
        class_clrs    = chikv_clrs,
        ag_clrs       = c("lightblue", "white", "#D7301F"),
        pt_size       = 0.75,
        circle_cells  = FALSE,
        show_scale_title = TRUE
      )
  })
```

```{r "CHIKV Ag class bargraphs"}
# Format data for bargraphs
chikv_bar_dat <- chikv_dat %>%
  filter(subtype %in% rf_typs) %>%
  group_by(treatment, rep, subtype) %>%
  summarize(
    frac_high = sum(pred_grp == "Ag-competent") / n(),
    .groups = "drop"
  ) %>%
  mutate(subtype = fct_relevel(subtype, rf_typs))

# Create Ag class bargraphs
chikv_ag_bars <- chikv_bar_dat %>%
  filter(subtype == rf_cell_type) %>%
  create_chikv_class_bars(ttl = str_c("fraction Ag-competent ", rf_cell_type, "s")) +
  theme(strip.text = element_blank())

# # Create final panels
# mid <- plot_grid(
#   chikv_u$cLEC[[2]], chikv_ag_bars,
#   align      = "h",
#   axis       = "tb",
#   rel_widths = c(1, 0.3), 
#   labels     = c("B", "C"),
#   label_size = ttl_pt2 * 1.5,
#   hjust      = 0.5,
#   vjust      = 1
# )
```

```{r "CHIKV Ag module score boxplots"}
# Format data for plotting
chikv_bx_dat <- chikv_dat %>%
  filter(subtype %in% rf_typs) %>%
  mutate(subtype = fct_relevel(subtype, rf_typs))

# Plot Ag module scores
chikv_ag_bxs <- chikv_bx_dat %>%
  filter(lec_subtype == rf_cell_type) %>%
  create_chikv_module_boxes(
    module_clmns = "cLEC_high",
    ttl          = "cLEC Ag-high module",
    n_p          = length(ag_modules)
  ) +
  theme(strip.text = element_blank())
  
# # Create final figure
# bot <- plot_grid(
#   chikv_u$cLEC[[1]], chikv_ag_bxs,
#   align = "h",
#   axis  = "tb",
#   rel_widths = c(1, 0.3),
#   labels     = c("D", "E"),
#   label_size = ttl_pt2 * 1.5,
#   hjust      = 0.5
# )
```

```{r "CHIKV Ag-high heatmap"}
# Ag-high heatmap data
heat_feats <- rf_feats$cLEC$high
heat_feats <- heat_feats[heat_feats %in% rownames(chikv_so)]

heat_dat <- chikv_so %>%
  FetchData(c("treatment", "rep", "lec_subtype", heat_feats)) %>%
  filter(lec_subtype %in% rf_cell_type) %>%
  pivot_longer(all_of(heat_feats)) %>%
  group_by(lec_subtype, treatment, rep, name) %>%
  summarize(value = mean(value), .groups = "drop")

gn_lvls <- heat_dat %>%
  pivot_wider(names_from = treatment, values_from = value) %>%
  mutate(value = CHIKV / mock, .keep = "unused") %>%
  arrange(desc(value)) %>%
  # pivot_wider(names_from = lec_subtype, values_from = value) %>%
  # arrange(desc(!!sym(rf_typs[1])), desc(!!sym(rf_typs[2])), desc(!!sym(rf_typs[3]))) %>%
  pull(name) %>%
  unique()
  
heat_dat <- heat_dat %>%
  group_by(name) %>%
  mutate(value = as.numeric(scale(value))) %>%
  ungroup() %>%
  mutate(
    treatment   = fct_relevel(treatment, c("mock", "CHIKV")),
    name        = fct_relevel(name, rev(gn_lvls))
    # lec_subtype = fct_relevel(lec_subtype, rev(rf_typs))
  )

# Ag heatmap
heat <- heat_dat %>%
  ggplot(aes(name, rep, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = c("#56B4E9", "white", "#D7301F")) +
  facet_wrap(~ treatment, ncol = 1, strip.position = "left") +
  guides(fill = guide_colorbar(
    barwidth = unit(150, "pt"),
    barheight = unit(6, "pt")
  )) +
  base_theme +
  theme(
    aspect.ratio    = length(rf_typs) / length(heat_feats),
    strip.placement = "outside",
    strip.clip      = "off",
    legend.position = "bottom",
    legend.title    = element_blank(),
    axis.title      = element_blank(),
    axis.text.x     = element_text(angle = 90, hjust = 1, vjust = 0.5)
  )
```

```{r "Fig 5", fig.width = 12, fig.height = 17}
# Combine and annotate center panels
c("D", "B") %>%
  iwalk(~ {
    chikv_u$cLEC[[.y]][[1]] <<- chikv_u$cLEC[[.y]][[1]] + labs(tag = .x)
  })

mid <- wrap_plots(
  chikv_u$cLEC[[2]],
  chikv_ag_bars + labs(tag = "C"),
  chikv_u$cLEC[[1]],
  chikv_ag_bxs + labs(tag = "E"),
  widths = c(1, 0.3)
) &
  theme(plot.tag = element_text(size = ttl_pt2 * 1.5, face = "bold"))

# Final figure
fig5 <- plot_grid(
  top, mid, heat,
  ncol       = 1,
  labels     = c("A", "", "F"),
  label_size = ttl_pt2 * 1.5,
  hjust      = c(0, 0, -0.5),
  vjust      = c(1.5, 1, 0),
  rel_heights = c(1, 2.5, 1)
)

fig5


# # Save figure for Tem's grant
# c("C", "A") %>%
#   iwalk(~ {
#     chikv_u$cLEC[[.y]][[1]] <<- chikv_u$cLEC[[.y]][[1]] + labs(tag = .x)
#   })
# 
# mid <- wrap_plots(
#   chikv_u$cLEC[[2]],
#   chikv_ag_bars + labs(tag = "B"),
#   chikv_u$cLEC[[1]],
#   chikv_ag_bxs + labs(tag = "D"),
#   widths = c(1, 0.3)
# ) &
#   theme(plot.tag = element_text(size = ttl_pt2 * 1.5, face = "bold"))
# 
# new_fig5 <- plot_grid(
#   mid, heat,
#   ncol       = 1,
#   labels     = c("", "E"),
#   label_size = ttl_pt2 * 1.5,
#   hjust      = c(0, -0.5),
#   vjust      = c(1, 0),
#   rel_heights = c(2.5, 1)
# )
# 
# new_fig5 %>%
#   ggsave(
#     filename = here("results", str_c(Sys.Date(), "_fig5.png")),
#     device = "png",
#     width  = 12,
#     height = 13.2,
#     bg     = "white",
#     dpi    = 600
#   )
```


```{r "OLD CHIKV Ag module score UMAPs", eval = FALSE}
# cLEC Ag-high module scores
ag_score_u <- chikv_dat %>%
  plot_scatter(
    "cLEC_high",
    x           = "hUMAP_1",
    y           = "hUMAP_2",
    group_col   = "treatment",
    plot_colors = c("#56B4E9", "white", "#D7301F"),
    outline     = TRUE,
    # size        = 0.5,
    size        = 0.75,
    stroke      = 0.7,
    n_label     = "none"
  ) +
  guides(fill = guide_colorbar(
    title = "cLEC\nAg-high module",
    title.hjust = 0.5,
    barwidth = unit(150, "pt"),
    barheight = unit(10, "pt")
  )) +
  chikv_u_theme +
  theme(
    legend.position = "bottom"
    # legend.title = element_blank()
  )
```

<br>

<br>

### Figure S6

A.  The fraction of cells predicted to be Ag-competent is shown for mock and
    CHIKV-infected mice for each biological replicate.
B.  Ag-high and Ag-low module scores are shown for mock and CHIKV-infected mice
    for each biological replicate. P values were calculated using a two-sided
    Wilcoxon rank sum test with Bonferroni correction.

```{r "Fig S6", fig.width = 6, fig.height = 9, out.width = "50%"}
# Bargraphs for all LEC subtypes
all_chikv_bars <- chikv_bar_dat %>%
  create_chikv_class_bars(ttl = "fraction Ag-competent")

# Boxplots for all Ag modules
all_chikv_bxs <- chikv_bx_dat %>%
  create_chikv_module_boxes(
    module_clmns = ag_modules,
    ttl = "module score"
  )

# Create final figure
figS6 <- plot_grid(
  all_chikv_bars, all_chikv_bxs,
  labels = c("A", "B"),
  label_size = ttl_pt2 * 1.5,
  ncol  = 1,
  align = "v",
  axis  = "rl",
  rel_heights = c(0.6, 1)
)

figS6
```

---

<br>

<br>

## Session info

```{r "session info"}
sessionInfo()
```



```{r "Pdpn Ag correlation", eval = FALSE}
# Format data for plotting
clmns <- c(
  "Pdpn", "Ag_score",
  "subtype", "mouse",
  "pred_grp", "Ag_class", "training_data"
)

dat <- lec_so %>%
  FetchData(clmns) %>%
  mutate(
    Ag_class = fct_relevel(Ag_class, c("low", "high")),
    pred_grp = fct_relevel(pred_grp, names(pred_clrs)),
    subtype  = fct_relevel(subtype, rf_typs)
  ) %>%
  filter(!training_data)
  
# Plot Pdpn for Ag-class
dat %>%
  filter(subtype %in% rf_typs, mouse != "d2") %>%
  ggplot(aes(subtype, Pdpn, fill = pred_grp)) +
  geom_boxplot(outlier.size = 0.5) +
  scale_fill_manual(values = pred_clrs) +
  facet_wrap(~ mouse) +
  base_theme +
  theme(
    aspect.ratio = 0.6,
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

ggsave(
  here("results/Pdpn_Ag_class.png"),
  dpi = 300
)

# Plot Pdpn for LEC subsets
dat %>%
  plot_violin(
    "Pdpn",
    "subtype",
    "mouse",
    method = "boxplot",
    plot_colors = lec_clrs,
    color = "black",
    alpha = 1,
    outlier.size = 0.5
  ) +  
  base_theme +
  theme(
    aspect.ratio = 0.6,
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

ggsave(
  here("results/Pdpn_LEC_subsets.png"),
  dpi = 300
)

# Plot Ag-score for LEC subsets
dat %>%
  plot_violin(
    "Ag_score",
    "subtype",
    "mouse",
    method = "boxplot",
    plot_colors = lec_clrs,
    color = "black",
    alpha = 1,
    outlier.size = 0.5
  ) +  
  base_theme +
  theme(
    aspect.ratio = 0.6,
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

ggsave(
  here("results/Ag-score_LEC_subsets.png"),
  dpi = 300
)

# Scatter plot comparing Pdpn expression and Ag-score
dat %>%
  filter(Pdpn > 0, Ag_score > 0) %>%
  ggplot(aes(Ag_score, Pdpn, color = subtype)) +
  geom_point() +
  geom_smooth(method = "lm", color = "black", linetype = 2) +
  scale_color_manual(values = lec_clrs) +
  base_theme +
  theme(
    aspect.ratio = 1
  )

ggsave(
  here("results/Pdpn_Ag-score_scatter.png"),
  dpi = 300
)
```

```{r "RF TEST 1", eval = FALSE}
# GOALS
# 1. train model for LEC subsets (~80% accuracy for cLECs)
# 2. train general model for all LECs (~60-90% for all subsets)

# Identify Ag markers for all LECs
degs_1 <- lec_so %>%
  wilcoxauc(group_by = "Ag_class") %>%
  filter(
    padj < 0.05, logFC > 0.25,
    !grepl("^(mt-|R[pls])", feature)
  ) %>%
  arrange(padj)

# Identify Ag markers for subtype
degs_2 <- unique(lec_so$subtype) %>%
  map_dfr(~ {
    res <- lec_so %>%
      subset(subtype == .x) %>%
      wilcoxauc(group_by = "Ag_class")

    if (nrow(res) == 0) return(NULL)

    res %>%
      filter(
        padj < 0.05, logFC > 0.25,
        !grepl("^(mt-|R[pls])", feature)
      ) %>%
      mutate(subtype = .x) %>%
      arrange(padj)
  })

# Identify markers for subtype and mouse
degs_3 <- lec_so@meta.data %>%
  distinct(subtype, mouse) %>%
  pmap_dfr(~ {
    res <- lec_so %>%
      subset(subtype == ..1 & mouse == ..2)
    
    if (n_distinct(res$Ag_class) < 2) return(NULL)
    
    res <- res %>%
      wilcoxauc(group_by = "Ag_class")

    if (nrow(res) == 0) return(NULL)

    res %>%
      filter(
        padj < 0.05, logFC > 0.25,
        !grepl("^(mt-|R[pls])", feature)
      ) %>%
      mutate(subtype = ..1, mouse = ..2) %>%
      arrange(padj)
  })

degs <- bind_rows(degs_1, degs_2, degs_3) %>%
  distinct(feature, group)

# Format RF data
# cat_feats <- c("tm", "subtype", "Ag_class")
# dat_col   <- "Ag_score"
cat_feats <- c("tm", "mouse", "subtype")
feats     <- unique(degs$feature)
dat_col   <- "Ag_class"

rf_dat <- lec_so %>%
  FetchData(c(dat_col, cat_feats, feats)) %>%
  mutate(!!sym(dat_col) := fct_relevel(!!sym(dat_col), "high", "low")) %>%
  rename_with(~ str_replace_all(.x, "-", "_")) %>%
  rename_with(~ str_replace_all(.x, "\\(|\\)", "_")) %>%
  rename_with(~ str_c("x", .x), matches("^[0-9]"))

set.seed(42)

rf_split <- rf_dat %>%
  initial_split(prop = 0.1)

rf_test <- testing(rf_split)

rf_train <- training(rf_split) %>%
  dplyr::select(-all_of(cat_feats))

# Tune model
tune_rf <- function(dat_df, dat_col, param_lst, return_best = TRUE) {
  
  rf_params <- expand.grid(param_lst)
  
  res <- pmap_dfr(rf_params, ~ {
    res <- ranger(
      as.formula(str_c(dat_col, " ~ .")),
      data       = dat_df,
      importance = "impurity",
      seed       = 42,
      ...
    )
    
    err <- sqrt(res$prediction.error)
    
    tibble(..., oob_rmse = err)
  })
  
  if (return_best) {
    res <- res %>%
      dplyr::filter(oob_rmse == min(oob_rmse)) %>%
      dplyr::select(-oob_rmse) %>%
      head(1) %>%
      as.list()
  }
  
  res
}

param_lst <- list(
  mtry            = seq(1, ncol(rf_train), 50),
  min.node.size   = c(1:3, 5, 10)
)

rf_params <- rf_train %>%
  tune_rf(
    dat_col   = "Ag_class",
    param_lst = param_lst
  )

# Fit RF
rf_mod <- pmap(rf_params, ~ {
  ranger(
    Ag_class ~ .,
    data       = rf_train,
    importance = "impurity",
    ...
  )
}) %>%
  pluck(1)

# Get predictions
prds <- rf_mod %>%
  predict(rf_test) %>%
  predictions()

rf_res <- rf_test %>%
  as_tibble(rownames = ".cell_id") %>%
  dplyr::select(.cell_id, subtype, tm, mouse, Ag_class) %>%
  mutate(
    prediction = prds,
    correct    = Ag_class == prediction
  ) %>%
  group_by(subtype, mouse) %>%
  summarize(
    n       = n(),
    correct = sum(correct) / n,
    .groups = "drop"
  ) %>%
  arrange(desc(correct))

rf_res %>%
  mutate(lab = as.character(round(correct, 2))) %>%
  ggplot(aes(subtype, mouse, fill = correct)) +
  geom_tile() +
  geom_text(aes(label = lab)) +
  scale_fill_gradientn(colours = c("lightblue", "red")) +
  djvdj_theme()

# Get top regression variables
rf_marks <- tibble(
  gene   = names(rf_mod$variable.importance),
  impact = rf_mod$variable.importance
) %>%
  mutate(
    rank = row_number(desc(impact)),
    gene = ifelse(gene %in% cat_feats, gene, str_replace(gene, "_", "-"))
  ) %>%
  arrange(rank) %>%

  left_join(degs, by = c(gene = "feature"))

# # Create boxplots
# n_gns <- 50
# 
# plt_gns <- rf_marks %>%
#   filter(!gene %in% c("Ag-score", "Ag_class", "orig.ident", "tm")) %>%
#   pull(gene) %>%
#   head(n_gns) %>%
#   c("Ag_score", .)
# 
# plt_dat <- rf_so %>%
#   subset(subtype == typ) %>%
#   FetchData(c("tm", "Ag_class", plt_gns)) %>%
#   pivot_longer(all_of(plt_gns)) %>%
#   mutate(
#     tm       = fct_relevel(as.character(tm), tms),
#     Ag_class = fct_relevel(Ag_class, c("low", "high")),
#     name     = fct_relevel(name, plt_gns)
#   )
#   
# plt_dat %>%
#   ggplot(aes(tm, value, fill = name, color = name, alpha = Ag_class)) +
#   geom_boxplot() +
#   facet_wrap(~ name, scales = "free_y") +
#   scale_alpha_manual(values = c(low = 0.25, high = 0.75)) +
#   djvdj_theme()
# 
# # Create scatter plots
# plt_dat <- rf_so %>%
#   subset(subtype == typ) %>%
#   FetchData(c("tm", "Ag_class", plt_gns)) %>%
#   pivot_longer(all_of(plt_gns[-1])) %>%
#   mutate(
#     tm       = fct_relevel(as.character(tm), tms),
#     Ag_class = fct_relevel(Ag_class, c("low", "high")),
#     name     = fct_relevel(name, plt_gns)
#   )
# 
# plt_dat %>%
#   ggplot(aes(Ag_score, value, color = Ag_class)) +
#   geom_point(size = 0.1) +
#   facet_wrap(~ name, scales = "free") +
#   djvdj_theme() +
#   theme(aspect.ratio = 1)
```

```{r "RF TEST 2", eval = FALSE}
# GOALS
# 1. train model for LEC subsets for Ag groups (1-4)
# 2. identify cells in mock that are predicted to take up Ag

# Identify Ag markers for all LECs for each mouse
deg_so <- lec_so %>%
  SplitObject("mouse")

plan(multisession, workers = min(length(deg_so), 8))

degs_1 <- deg_so %>%
  future_imap_dfr(~ {
    res <- .x %>%
      wilcoxauc(group_by = "Ag_group")

    if (nrow(res) == 0) return(NULL)

    res %>%
      filter(
        padj < 0.05, logFC > 0.25,
        !grepl("^(mt-|R[pls])", feature)
      ) %>%
      mutate(mouse = .y) %>%
      arrange(padj)
  })

# Identify Ag markers for subtype
deg_so <- lec_so %>%
  SplitObject("subtype")

plan(multisession, workers = min(length(deg_so), 8))

degs_2 <- deg_so %>%
  future_imap_dfr(~ {
    res <- .x %>%
      wilcoxauc(group_by = "Ag_group")

    if (nrow(res) == 0) return(NULL)

    res %>%
      filter(
        padj < 0.05, logFC > 0.25,
        !grepl("^(mt-|R[pls])", feature)
      ) %>%
      mutate(subtype = .y) %>%
      arrange(padj)
  })

rm(deg_so)

# Identify markers for subtype and mouse
degs_3 <- lec_so@meta.data %>%
  distinct(subtype, mouse) %>%
  pmap_dfr(~ {
    res <- lec_so %>%
      subset(subtype == ..1 & mouse == ..2)
    
    if (n_distinct(res$Ag_class) < 2) return(NULL)
    
    res <- res %>%
      wilcoxauc(group_by = "Ag_group")

    if (nrow(res) == 0) return(NULL)

    res %>%
      filter(
        padj < 0.05, logFC > 0.25,
        !grepl("^(mt-|R[pls])", feature)
      ) %>%
      mutate(subtype = ..1, mouse = ..2) %>%
      arrange(padj)
  })

degs <- bind_rows(degs_1, degs_2, degs_3) %>%
  distinct(feature, group)

# Format RF data
dat_col   <- "Ag_group"
cat_feats <- c("tm", "mouse", "subtype")
feats     <- unique(degs$feature)

rf_dat <- lec_so %>%
  FetchData(c(dat_col, cat_feats, feats)) %>%
  rename_with(~ str_replace_all(.x, "-", "_")) %>%
  rename_with(~ str_replace_all(.x, "\\(|\\)", "_")) %>%
  rename_with(~ str_c("x", .x), matches("^[0-9]"))

set.seed(42)

rf_split <- rf_dat %>%
  initial_split(prop = 0.1)

rf_test <- testing(rf_split)

rf_train <- training(rf_split) %>%
  dplyr::select(-all_of(cat_feats))

# Tune model
tune_rf <- function(dat_df, dat_col, param_lst, return_best = TRUE) {
  
  rf_params <- expand.grid(param_lst)
  
  res <- pmap_dfr(rf_params, ~ {
    res <- ranger(
      as.formula(str_c(dat_col, " ~ .")),
      data       = dat_df,
      importance = "impurity",
      seed       = 42,
      ...
    )
    
    err <- sqrt(res$prediction.error)
    
    tibble(..., oob_rmse = err)
  })
  
  if (return_best) {
    res <- res %>%
      dplyr::filter(oob_rmse == min(oob_rmse)) %>%
      dplyr::select(-oob_rmse) %>%
      head(1) %>%
      as.list()
  }
  
  res
}

param_lst <- list(
  mtry            = seq(1, ncol(rf_train), 100),
  min.node.size   = c(1:3, 5, 10)
)

rf_params <- rf_train %>%
  tune_rf(
    dat_col   = "Ag_group",
    param_lst = param_lst
  )

# Fit model
rf_mod <- pmap(rf_params, ~ {
  ranger(
    Ag_class ~ .,
    data       = rf_train,
    importance = "impurity",
    ...
  )
}) %>%
  pluck(1)

# Get predictions
prds <- rf_mod %>%
  predict(rf_test) %>%
  predictions()

rf_res <- rf_test %>%
  as_tibble(rownames = ".cell_id") %>%
  dplyr::select(.cell_id, subtype, tm, mouse, Ag_class) %>%
  mutate(
    prediction = prds,
    correct    = Ag_class == prediction
  ) %>%
  group_by(subtype, mouse) %>%
  summarize(
    n       = n(),
    correct = sum(correct) / n,
    .groups = "drop"
  ) %>%
  arrange(desc(correct))

rf_res %>%
  mutate(lab = as.character(round(correct, 2))) %>%
  ggplot(aes(subtype, mouse, fill = correct)) +
  geom_tile() +
  geom_text(aes(label = lab)) +
  scale_fill_gradientn(colours = c("lightblue", "red")) +
  djvdj_theme()

# Get top regression variables
rf_marks <- tibble(
  gene   = names(rf_mod$variable.importance),
  impact = rf_mod$variable.importance
) %>%
  mutate(
    rank = row_number(desc(impact)),
    gene = ifelse(gene %in% cat_feats, gene, str_replace(gene, "_", "-"))
  ) %>%
  arrange(rank) %>%

  left_join(degs, by = c(gene = "feature"))



# # Create boxplots
# n_gns <- 50
# 
# plt_gns <- rf_marks %>%
#   filter(!gene %in% c("Ag-score", "Ag_class", "orig.ident", "tm")) %>%
#   pull(gene) %>%
#   head(n_gns) %>%
#   c("Ag_score", .)
# 
# plt_dat <- rf_so %>%
#   subset(subtype == typ) %>%
#   FetchData(c("tm", "Ag_class", plt_gns)) %>%
#   pivot_longer(all_of(plt_gns)) %>%
#   mutate(
#     tm       = fct_relevel(as.character(tm), tms),
#     Ag_class = fct_relevel(Ag_class, c("low", "high")),
#     name     = fct_relevel(name, plt_gns)
#   )
#   
# plt_dat %>%
#   ggplot(aes(tm, value, fill = name, color = name, alpha = Ag_class)) +
#   geom_boxplot() +
#   facet_wrap(~ name, scales = "free_y") +
#   scale_alpha_manual(values = c(low = 0.25, high = 0.75)) +
#   djvdj_theme()
# 
# # Create scatter plots
# plt_dat <- rf_so %>%
#   subset(subtype == typ) %>%
#   FetchData(c("tm", "Ag_class", plt_gns)) %>%
#   pivot_longer(all_of(plt_gns[-1])) %>%
#   mutate(
#     tm       = fct_relevel(as.character(tm), tms),
#     Ag_class = fct_relevel(Ag_class, c("low", "high")),
#     name     = fct_relevel(name, plt_gns)
#   )
# 
# plt_dat %>%
#   ggplot(aes(Ag_score, value, color = Ag_class)) +
#   geom_point(size = 0.1) +
#   facet_wrap(~ name, scales = "free") +
#   djvdj_theme() +
#   theme(aspect.ratio = 1)
```

```{r "RF TEST 3", eval = FALSE}
# GOALS
# 1. train model for LEC subsets for Ag groups (1-4)
# 2. identify cells in mock that are predicted to take up Ag

# Identify DEGs ----
# Identify Ag markers for all LECs for each mouse
grp <- "Ag_class"
fc  <- 0

deg_so <- lec_so %>%
  SplitObject("mouse")

degs_1 <- deg_so %>%
  imap_dfr(~ {
    res <- .x %>%
      wilcoxauc(group_by = grp)

    if (nrow(res) == 0) return(NULL)

    res %>%
      mutate(subtype = "all", mouse = .y) %>%
      arrange(padj)
  })

# Identify Ag markers for subtype
deg_so <- lec_so %>%
  SplitObject("subtype")

degs_2 <- deg_so %>%
  imap_dfr(~ {
    res <- .x %>%
      wilcoxauc(group_by = grp)

    if (nrow(res) == 0) return(NULL)

    res %>%
      mutate(subtype = .y, mouse = "all") %>%
      arrange(padj)
  })

rm(deg_so)

# Identify markers for subtype and mouse
degs_3 <- lec_so@meta.data %>%
  distinct(subtype, mouse) %>%
  pmap_dfr(~ {
    res <- lec_so %>%
      subset(subtype == ..1 & mouse == ..2)
    
    if (n_distinct(res$Ag_class) < 2) return(NULL)
    
    res <- res %>%
      wilcoxauc(group_by = grp)

    if (nrow(res) == 0) return(NULL)

    res %>%
      mutate(subtype = ..1, mouse = ..2) %>%
      arrange(padj)
  })

degs <- bind_rows(degs_1, degs_2, degs_3) %>%
  filter(
    padj < 0.05, logFC > fc,
    !grepl("^(mt-|R[pls])", feature)
  )

# Format data ----
# dat_col    <- "Ag_group"
dat_col    <- "Ag_class_2"
feats      <- unique(degs$feature)
other_vars <- NULL

feats <- degs %>%
  filter(subtype == "cLEC", mouse == "d14") %>%
  pull(feature) %>%
  unique()

other_info <- c("tm", "subtype", "mouse")

rf_dat <- lec_so %>%
  subset(subtype == "cLEC" & mouse == "d14") %>%

  FetchData(unique(c(dat_col, other_info, other_vars, feats))) %>%
  
  # mutate(across(all_of(feats), ~ as.numeric(scale(.x)))) %>%
  # mutate(!!sym(dat_col) := factor(!!sym(dat_col))) %>%
  # mutate(
  #   across(all_of(c(dat_col, other_vars)), ~ as.numeric(factor(.x)))
  # ) %>%
  
  rename_with(~ str_replace_all(.x, "-", "_")) %>%
  rename_with(~ str_replace_all(.x, "\\(|\\)", "_")) %>%
  rename_with(~ str_c("x", .x), matches("^[0-9]"))

# Adjust groups so they are balanced
sam_sz <- rf_dat %>%
  group_by(!!sym(dat_col), mouse) %>%
  summarize(n = n(), .groups = "drop") %>%
  pull(n)

# sam_sz <- max(100, min(sam_sz))
sam_sz <- 135

# Split into train/test data
# rf_split <- initial_split(rf_dat, prop = 0.5)
# rf_train <- training(rf_split)
# rf_test  <- testing(rf_split)

rf_train <- rf_dat %>%
  rownames_to_column(".cell_id") %>%
  group_by(!!sym(dat_col), mouse) %>%
  filter(n() >= sam_sz) %>%
  dplyr::sample_n(sam_sz) %>%
  ungroup() %>%
  dplyr::select(-all_of(other_info[!other_info %in% other_vars])) %>%
  column_to_rownames(".cell_id")

rf_test <- rf_dat[!rownames(rf_dat) %in% rownames(rf_train), ]

# Tune RF model ----
tune_rf <- function(dat_df, dat_col, param_lst, return_best = TRUE) {
  
  rf_params <- expand.grid(param_lst)
  
  res <- pmap_dfr(rf_params, ~ {
    res <- ranger(
      as.formula(str_c("factor(", dat_col, ") ~ .")),
      data       = dat_df,
      importance = "impurity",
      seed       = 42,
      ...
    )
    
    err <- sqrt(res$prediction.error)
    
    tibble(..., oob_rmse = err)
  })
  
  if (return_best) {
    res <- res %>%
      dplyr::filter(oob_rmse == min(oob_rmse)) %>%
      dplyr::select(-oob_rmse) %>%
      head(1) %>%
      as.list()
  }
  
  res
}

param_lst <- list(
  # mtry          = c(10, 50, 200, 300),
  mtry          = seq(10, 60, 10),
  min.node.size = c(1, 2, 5),
  num.trees     = 1000
)

rf_params <- rf_train %>%
  tune_rf(
    dat_col   = dat_col,
    param_lst = param_lst
  )

# rf_params <- list(
#   mtry = 10,
#   min.node.size = 1,
#   num.trees = 1000
#   # max.depth = 5
# )

# Fit RF model
rf_mod <- pmap(rf_params, ~ {
  ranger(
    as.formula(str_c("factor(", dat_col, ") ~ .")),
    data       = rf_train,
    importance = "impurity",
    ...
  )
}) %>%
  pluck(1)

# Identify top features and re-fit
top_feats <- importance(rf_mod) %>%
  sort(decreasing = TRUE)

top_feats <- top_feats[top_feats > 0]

rf_params$mtry <- min(length(top_feats), rf_params$mtry)

rf_mod_top <- pmap(rf_params, ~ {
  ranger(
    as.formula(str_c("factor(", dat_col, ") ~ .")),
    data       = rf_train[, c(dat_col, names(top_feats))],
    importance = "impurity",
    ...
  )
}) %>%
  pluck(1)

# prds <- rf_mod_top %>%
prds <- rf_mod %>%
  predict(rf_test) %>%
  predictions()

# Try logit regression ----
logit_mod <- rf_train %>%
  mutate(
    Ag_class = ifelse(Ag_class == "high", 1, 0),
    Ag_class = factor(Ag_class)
  ) %>%
  glm(
    data = .,
    # as.factor(Ag_class) ~ Cav1 + Marcks | as.numeric(as.factor(mouse)),
    # as.factor(Ag_class) ~ Cav1 + Marcks,
    Ag_class ~ .,
    control = glm.control(maxit = 200),
    family = binomial
  )

prob <- logit_mod %>%
  predict(newdata = rf_test, type = "response")

prds <- vector("character", length(prob))
prds[prob > 0.7] <- "high"
prds[prob <= 0.7] <- "low"

# Try glmer ----
# library(lme4)
# 
# logit_mod <- rf_train %>%
#   mutate(
#     Ag_class = ifelse(Ag_class == "high", 1, 0),
#     Ag_class = factor(Ag_class)
#   ) %>%
#   glmer(
#     data = .,
#     # as.factor(Ag_class) ~ Cav1 + Marcks | as.numeric(as.factor(mouse)),
#     # as.factor(Ag_class) ~ Cav1 + Marcks,
#     # Ag_class ~ (. | factor(mouse)),
#     # control = glm.control(maxit = 200),
#     Ag_class ~ (. | factor(exp)),
#     family = binomial
#   )
# 
# prob <- logit_mod %>%
#   predict(newdata = rf_test, type = "response")
# 
# prds <- vector("character", length(prob))
# prds[prob > 0.5] <- "high"
# prds[prob <= 0.5] <- "low"


# Get predictions ----
rf_res <- rf_test %>%
  as_tibble(rownames = ".cell_id") %>%
  dplyr::select(.cell_id, subtype, all_of(c(dat_col, other_vars, other_info))) %>%
  mutate(
    prediction = prds,
    correct    = !!sym(dat_col) == prediction
  ) %>%
  # group_by(subtype, !!sym(other_vars)) %>%
  group_by(!!!syms(c(other_info, dat_col))) %>%
  summarize(
    n       = n(),
    correct = sum(correct) / n,
    .groups = "drop"
  ) %>%
  arrange(desc(correct))

# Plot results
rf_res %>%
  mutate(lab = as.character(round(correct, 2))) %>%
  ggplot(aes(subtype, mouse, fill = correct)) +
  geom_tile() +
  geom_text(aes(label = lab)) +
  facet_wrap(~ Ag_class_2) +
  scale_fill_gradientn(colours = c("lightblue", "red")) +
  djvdj_theme()

# Get top regression variables
rf_marks <- tibble(
  gene   = names(rf_mod$variable.importance),
  impact = rf_mod$variable.importance
) %>%
  mutate(
    rank = row_number(desc(impact)),
    gene = ifelse(gene %in% cat_feats, gene, str_replace(gene, "_", "-"))
  ) %>%
  arrange(rank) %>%

  left_join(degs, by = c(gene = "feature"))



# # Create boxplots
# n_gns <- 50
# 
# plt_gns <- rf_marks %>%
#   filter(!gene %in% c("Ag-score", "Ag_class", "orig.ident", "tm")) %>%
#   pull(gene) %>%
#   head(n_gns) %>%
#   c("Ag_score", .)
# 
# plt_dat <- rf_so %>%
#   subset(subtype == typ) %>%
#   FetchData(c("tm", "Ag_class", plt_gns)) %>%
#   pivot_longer(all_of(plt_gns)) %>%
#   mutate(
#     tm       = fct_relevel(as.character(tm), tms),
#     Ag_class = fct_relevel(Ag_class, c("low", "high")),
#     name     = fct_relevel(name, plt_gns)
#   )
#   
# plt_dat %>%
#   ggplot(aes(tm, value, fill = name, color = name, alpha = Ag_class)) +
#   geom_boxplot() +
#   facet_wrap(~ name, scales = "free_y") +
#   scale_alpha_manual(values = c(low = 0.25, high = 0.75)) +
#   djvdj_theme()
# 
# # Create scatter plots
# plt_dat <- rf_so %>%
#   subset(subtype == typ) %>%
#   FetchData(c("tm", "Ag_class", plt_gns)) %>%
#   pivot_longer(all_of(plt_gns[-1])) %>%
#   mutate(
#     tm       = fct_relevel(as.character(tm), tms),
#     Ag_class = fct_relevel(Ag_class, c("low", "high")),
#     name     = fct_relevel(name, plt_gns)
#   )
# 
# plt_dat %>%
#   ggplot(aes(Ag_score, value, color = Ag_class)) +
#   geom_point(size = 0.1) +
#   facet_wrap(~ name, scales = "free") +
#   djvdj_theme() +
#   theme(aspect.ratio = 1)
```

```{r "RF d14 TEST", eval = FALSE}
# GOALS
# 1. train model for LEC subsets for Ag groups (1-4)
# 2. identify cells in mock that are predicted to take up Ag

# Identify DEGs ----
# Identify Ag markers for all LECs for each mouse
grp <- "Ag_class_2"
fc  <- 0

deg_so <- lec_so %>%
  SplitObject("mouse")

degs_1 <- deg_so %>%
  imap_dfr(~ {
    res <- .x %>%
      wilcoxauc(group_by = grp)

    if (nrow(res) == 0) return(NULL)

    res %>%
      mutate(subtype = "all", mouse = .y) %>%
      arrange(padj)
  })

# Identify Ag markers for subtype
deg_so <- lec_so %>%
  SplitObject("subtype")

degs_2 <- deg_so %>%
  imap_dfr(~ {
    res <- .x %>%
      wilcoxauc(group_by = grp)

    if (nrow(res) == 0) return(NULL)

    res %>%
      mutate(subtype = .y, mouse = "all") %>%
      arrange(padj)
  })

rm(deg_so)

# Identify markers for subtype and mouse
degs_3 <- lec_so@meta.data %>%
  distinct(subtype, mouse) %>%
  pmap_dfr(~ {
    res <- lec_so %>%
      subset(subtype == ..1 & mouse == ..2)
    
    if (n_distinct(res[[grp]]) < 2) return(NULL)
    
    res <- res %>%
      wilcoxauc(group_by = grp)

    if (nrow(res) == 0) return(NULL)

    res %>%
      mutate(subtype = ..1, mouse = ..2) %>%
      arrange(padj)
  })

degs <- bind_rows(degs_1, degs_2, degs_3) %>%
  filter(
    padj < 0.05, logFC > fc,
    !grepl("^(mt-|R[pls])", feature),
    feature != "Malat1"
  )

degs %>%
  write_tsv(here(params$so_dir, "ag-high_degs_Ag_class_2.tsv.gz"))

# Most common degs
top_degs <- degs %>%
  group_by(subtype, feature, group) %>%
  summarize(n = n_distinct(mouse), .groups = "drop") %>%
  arrange(desc(n)) %>%
  filter(n > 1) %>%
  filter(!subtype %in% c("all", "BEC", "unassigned")) %>%
  group_by(feature, group) %>%
  summarize(n_typ = n_distinct(subtype), .groups = "drop") %>%
  arrange(desc(n_typ)) %>%
  filter(n_typ > 1)

# Format data ----
dat_col    <- "Ag_class_2"
feats      <- unique(degs$feature)
other_vars <- NULL

# feats <- degs %>%
#   filter(subtype == "cLEC", mouse == "d14") %>%
#   pull(feature) %>%
#   unique()
feats <- unique(top_degs$feature)

other_info <- c("tm", "subtype", "mouse")

rf_dat <- rf_so %>%
  # subset(subtype != "BEC" & subtype == "cLEC") %>%
  subset(subtype == "cLEC" & mouse == "d14") %>%

  FetchData(unique(c(dat_col, other_info, other_vars, feats))) %>%
  
  rename_with(~ str_replace_all(.x, "-", "_")) %>%
  rename_with(~ str_replace_all(.x, "\\(|\\)", "_")) %>%
  rename_with(~ str_c("x", .x), matches("^[0-9]"))

# Adjust groups so they are balanced
# Split into train/test data
sam_sz <- 135

rf_train <- rf_dat %>%
  rownames_to_column(".cell_id") %>%
  group_by(!!sym(dat_col), subtype, mouse) %>%
  filter(n() >= sam_sz) %>%
  dplyr::sample_n(sam_sz) %>%
  ungroup() %>%
  dplyr::select(-all_of(other_info[!other_info %in% other_vars])) %>%
  column_to_rownames(".cell_id")

rf_test <- rf_dat[!rownames(rf_dat) %in% rownames(rf_train), ]

# Tune RF model ----
tune_rf <- function(dat_df, dat_col, param_lst, return_best = TRUE) {
  
  rf_params <- expand.grid(param_lst)
  
  res <- pmap_dfr(rf_params, ~ {
    res <- ranger(
      as.formula(str_c("factor(", dat_col, ") ~ .")),
      data       = dat_df,
      importance = "impurity",
      seed       = 42,
      ...
    )
    
    err <- sqrt(res$prediction.error)
    
    tibble(..., oob_rmse = err)
  })
  
  if (return_best) {
    res <- res %>%
      dplyr::filter(oob_rmse == min(oob_rmse)) %>%
      dplyr::select(-oob_rmse) %>%
      head(1) %>%
      as.list()
  }
  
  res
}

param_lst <- list(
  mtry          = seq(2, length(feats), 5),
  min.node.size = c(1, 2, 5),
  num.trees     = 1000
)

rf_params <- rf_train %>%
  tune_rf(
    dat_col   = dat_col,
    param_lst = param_lst
  )

# Fit RF model
rf_mod <- pmap(rf_params, ~ {
  ranger(
    as.formula(str_c("factor(", dat_col, ") ~ .")),
    data       = rf_train,
    importance = "impurity",
    ...
  )
}) %>%
  pluck(1)

# Identify top features and re-fit
top_feats <- importance(rf_mod) %>%
  sort(decreasing = TRUE)

top_feats <- top_feats[top_feats > 0]

rf_params$mtry <- min(length(top_feats), rf_params$mtry)

rf_mod_top <- pmap(rf_params, ~ {
  ranger(
    as.formula(str_c("factor(", dat_col, ") ~ .")),
    data       = rf_train[, c(dat_col, names(top_feats))],
    importance = "impurity",
    ...
  )
}) %>%
  pluck(1)

# prds <- rf_mod %>%
  # predict(rf_dat) %>%
prds <- rf_mod_top %>%
  predict(rf_test) %>%
  predictions()

# Get predictions ----
# rf_res <- rf_test %>%
rf_res <- rf_dat %>%
  as_tibble(rownames = ".cell_id") %>%
  dplyr::select(.cell_id, subtype, all_of(c(dat_col, other_vars, other_info))) %>%
  mutate(
    prediction = prds,
    correct    = !!sym(dat_col) == prediction
  ) %>%
  group_by(!!!syms(c(other_info, dat_col))) %>%
  summarize(
    n       = n(),
    correct = sum(correct) / n,
    .groups = "drop"
  ) %>%
  arrange(desc(correct))

# Plot results
rf_res %>%
  mutate(lab = as.character(round(correct, 2))) %>%
  ggplot(aes(subtype, mouse, fill = correct)) +
  geom_tile() +
  geom_text(aes(label = lab)) +
  facet_wrap(~ Ag_class_2) +
  scale_fill_gradientn(colours = c("lightblue", "red")) +
  djvdj_theme()

# Get top regression variables
rf_marks <- tibble(
  gene   = names(rf_mod$variable.importance),
  impact = rf_mod$variable.importance
) %>%
  mutate(
    rank = row_number(desc(impact)),
    gene = ifelse(gene %in% cat_feats, gene, str_replace(gene, "_", "-"))
  ) %>%
  arrange(rank) %>%

  left_join(degs, by = c(gene = "feature"))

# Create boxplots
n_gns <- 50

plt_gns <- rf_marks %>%
  filter(!gene %in% c("Ag_score", dat_col, "orig.ident", "tm")) %>%
  pull(gene) %>%
  unique() %>%
  head(n_gns) %>%
  c("Ag_score", .) %>%
  str_remove("^x")

plt_dat <- lec_so %>%
  subset(subtype == typ) %>%
  FetchData(c("tm", dat_col, plt_gns)) %>%
  pivot_longer(all_of(plt_gns)) %>%
  mutate(
    tm         = fct_relevel(as.character(tm), tms),
    Ag_class_2 = fct_relevel(Ag_class_2, c("low", "high")),
    name       = fct_relevel(name, plt_gns)
  )

plt_dat %>%
  ggplot(aes(tm, value, fill = name, color = name, alpha = Ag_class_2)) +
  geom_boxplot() +
  facet_wrap(~ name, scales = "free_y") +
  scale_alpha_manual(values = c(low = 0.25, high = 0.75)) +
  djvdj_theme()

# # Create scatter plots
# plt_dat <- rf_so %>%
#   subset(subtype == typ) %>%
#   FetchData(c("tm", "Ag_class", plt_gns)) %>%
#   pivot_longer(all_of(plt_gns[-1])) %>%
#   mutate(
#     tm       = fct_relevel(as.character(tm), tms),
#     Ag_class = fct_relevel(Ag_class, c("low", "high")),
#     name     = fct_relevel(name, plt_gns)
#   )
# 
# plt_dat %>%
#   ggplot(aes(Ag_score, value, color = Ag_class)) +
#   geom_point(size = 0.1) +
#   facet_wrap(~ name, scales = "free") +
#   djvdj_theme() +
#   theme(aspect.ratio = 1)
```



```{r "RF DEGs", eval = FALSE}
# GOALS
# 1. train model for LEC subsets for Ag groups (1-4)
# 2. identify cells in mock that are predicted to take up Ag

# Identify Ag markers for all LECs for each mouse
grp <- "Ag_class_2"
fc  <- 0

deg_so <- lec_so %>%
  SplitObject("mouse")

degs_1 <- deg_so %>%
  imap_dfr(~ {
    res <- .x %>%
      wilcoxauc(group_by = grp)

    if (nrow(res) == 0) return(NULL)

    res %>%
      mutate(subtype = "all", mouse = .y) %>%
      arrange(padj)
  })

# Identify Ag markers for subtype
deg_so <- lec_so %>%
  SplitObject("subtype")

degs_2 <- deg_so %>%
  imap_dfr(~ {
    res <- .x %>%
      wilcoxauc(group_by = grp)

    if (nrow(res) == 0) return(NULL)

    res %>%
      mutate(subtype = .y, mouse = "all") %>%
      arrange(padj)
  })

rm(deg_so)

# Identify markers for subtype and mouse
degs_3 <- lec_so@meta.data %>%
  distinct(subtype, mouse) %>%
  pmap_dfr(~ {
    res <- lec_so %>%
      subset(subtype == ..1 & mouse == ..2)
    
    if (n_distinct(res[[grp]]) < 2) return(NULL)
    
    res <- res %>%
      wilcoxauc(group_by = grp)

    if (nrow(res) == 0) return(NULL)

    res %>%
      mutate(subtype = ..1, mouse = ..2) %>%
      arrange(padj)
  })

degs <- bind_rows(degs_1, degs_2, degs_3) %>%
  filter(
    padj < 0.05, logFC > fc,
    !grepl("^(mt-|R[pls])", feature),
    feature != "Malat1"
  )

# degs %>%
#   write_tsv(here(params$so_dir, "ag-high_degs_Ag_class_2.tsv.gz"))

# # Most common degs
# top_degs <- degs %>%
#   group_by(subtype, feature, group) %>%
#   summarize(n = n_distinct(mouse), .groups = "drop") %>%
#   arrange(desc(n))
#   filter(n > 1)
```

```{r "RF FUNCTIONS", eval = FALSE}
# Function to predict
rf_predict <- function(mod, dat, data_clmn, return_stats = TRUE) {
  
  prds <- mod %>%
    predict(dat) %>%
    predictions()
  
  res <- dat %>%
    as_tibble(rownames = ".cell_id") %>%
    dplyr::select(.cell_id, all_of(data_clmn)) %>%
    mutate(
      pred = prds,
      correct = !!sym(data_clmn) == pred
    ) %>%
    group_by(!!sym(data_clmn)) %>%
    mutate(frac_correct = sum(correct) / n()) %>%
    ungroup()
  
  if (!return_stats) return(column_to_rownames(res, ".cell_id"))
  
  res <- res %>%
    distinct(!!sym(data_clmn), frac_correct) %>%
    pivot_wider(names_from = all_of(data_clmn), values_from = frac_correct)
  
  res
}

run_rf <- function(df_in, data_clmn, param_lst, sample_size = 135,
                   feat_p = 0.05, min_feats = 20, sample_info = NULL) {
  
  # Format train/test data
  rf_dat <- df_in %>%
    rename_with(~ str_replace_all(.x, "-", "_")) %>%
    rename_with(~ str_replace_all(.x, "\\(|\\)", "_")) %>%
    rename_with(~ str_c("x", .x), matches("^[0-9]"))
  
  # Adjust groups so they are balanced
  # Split into train/test data
  sam_sz <- min(sample_size, min(table(rf_dat[[data_clmn]])))
  min_sz <- 40
  
  if (sam_sz < min_sz) {
    cli::cli_abort("All groups must have at least {min_sz} cells")
    
  } else if (sam_sz < sample_size) {
    cli::cli_warn(
      "{sample_size} is too large for the number of cells,
       a sample size of {sam_sz} was used"
    )
  }
  
  rf_train <- rf_dat %>%
    rownames_to_column(".cell_id") %>%
    group_by(!!sym(data_clmn))
  
  set.seed(42)
  
  rf_train <- rf_train %>%
    dplyr::sample_n(sam_sz) %>%
    ungroup() %>%
    dplyr::select(-any_of(sample_info)) %>%
    column_to_rownames(".cell_id")
  
  rf_test <- rf_dat[!rownames(rf_dat) %in% rownames(rf_train), ]
  
  # Train models
  rf_params <- expand.grid(param_lst)
  
  res <- pmap_dfr(rf_params, ~ {
    args <- list(...)
    
    if (args$mtry > (ncol(rf_train) - 1)) return(NULL)
    
    args$formula    <- as.formula(str_c("factor(", data_clmn, ") ~ ."))
    args$data       <- rf_train
    args$importance <- "impurity_corrected"
    args$seed       <- 42
    
    set.seed(42)
    
    raw_mod <- lift_dl(ranger)(args)
    
    raw_err <- raw_mod$prediction.error

    # Feature selection
    key_feats <- raw_mod %>%
      importance_pvalues() %>%
      .[, "pvalue"] %>%
      sort()
    
    max_p <- key_feats %>%
      head(min_feats) %>%
      max()
    
    feat_p <- max(max_p, feat_p)
    
    key_feats <- key_feats[key_feats < feat_p] %>%
      names()
    
    n_feats <- length(key_feats)
    
    # Re-train model
    args$mtry       <- min(args$mtry, n_feats)
    args$data       <- rf_train[, c(data_clmn, key_feats)]
    args$importance <- "none"
    
    set.seed(42)
    
    mod <- lift_dl(ranger)(args)
    
    err <- mod$prediction.error
    
    # Test model    
    prds <- mod %>%
      rf_predict(
        dat = rf_test,
        data_clmn = data_clmn
      )
                                         # only used for continuous
    # err <- sqrt(res$prediction.error)  # response variables
    
    # Format output table
    tibble(
      raw_mod   = list(raw_mod),
      mod       = list(mod),
      feats     = list(key_feats),
      n_train   = nrow(rf_train),
      n_test    = nrow(rf_test),
      # train_bcs = list(rownames(rf_train)),
      # test_bcs  = list(rownames(rf_test)),
      ...,
      n_feats = length(key_feats),
      oob_raw = raw_err,
      oob     = err
    ) %>%
      bind_cols(prds)
  })
  
  res
}
```

```{r "RF MODELS", eval = FALSE}
# Train models
# use all cLEC features
dat_col <- "Ag_class_2"

feats <- degs %>%
  filter(subtype == "cLEC") %>%
  # filter(subtype == "cLEC", mouse == "d14") %>%
  pull(feature) %>%
  unique()

other_info <- c("subtype", "mouse")

rf_dat <- lec_so %>%
  subset(subtype == "cLEC" & mouse == "d14") %>%
  FetchData(unique(c(dat_col, other_info, feats)))

param_lst <- list(
  mtry          = seq(2, length(feats), 10),
  min.node.size = c(1:10),
  num.trees     = seq(50, 1000, 50)
)

rf_res <- rf_dat %>%
  run_rf(
    data_clmn   = dat_col,
    param_lst   = param_lst,
    sample_info = other_info,
    sample_size = 135
  ) %>%
  # arrange(desc(high), desc(low))
  arrange(desc(low), desc(high))

# Pull best model
best_mod <- rf_res %>%
  head(1) %>%
  pull(mod) %>%
  .[[1]]

best_feats <- rf_res %>%
  head(1) %>%
  pull(feats) %>%
  unlist()

# Predict for other timepoints
tm_dat <- lec_so %>%
  subset(subtype == "cLEC") %>%
  FetchData(unique(c(dat_col, other_info, feats))) %>%
  rename_with(~ str_replace_all(.x, "-", "_")) %>%
  rename_with(~ str_replace_all(.x, "\\(|\\)", "_")) %>%
  rename_with(~ str_c("x", .x), matches("^[0-9]")) %>%
  split(.$mouse)

preds <- tm_dat %>%
  imap_dfr(~ {
    best_mod %>%
      rf_predict(.x, data_clmn = dat_col, return_stats = FALSE) %>%
      mutate(mouse = .y)
  })

# Plot predicted classifications
pred_meta <- preds %>%
  dplyr::select(pred, correct)

lec_so <- lec_so %>%
  AddMetaData(pred_meta)

lec_so %>%
  plot_scatter(
    "pred", "hUMAP_1", "hUMAP_2",
    group_col = "mouse",
    size = 0.1,
    panel_nrow = 1
  ) +
  theme(aspect.ratio = 0.9)

# Fraction predicted in each mouse
plt_dat <- lec_so %>%
  subset(subtype == "cLEC") %>%
  AddModuleScore(features = best_feats, name = "Ag_module") %>%
  mutate_meta(
    mutate,
    pred_grp = case_when(
      Ag_class_2 == "high" ~ Ag_class_2,
      pred       == "high" ~ "high-predicted",
      TRUE                 ~ Ag_class_2
    ),
    mouse = fct_relevel(mouse, mice)
  )

plt_dat %>%
  plot_frequency("pred_grp", cluster_col = "mouse")

# Module scores for each prediction group
# expect that these genes will be highest expressed in high, high-predicted,
# lowest in low
# 3wk sample shows opposite
plt_dat@meta.data %>%
  # ggplot(aes(mouse, Ag_module1, fill = Ag_class_2)) +
  ggplot(aes(mouse, Ag_module1, fill = pred_grp)) +
  # geom_violin(scale = "width", draw_quantiles = c(0.25, 0.75, 0.5)) +
  geom_boxplot() +
  djvdj_theme()
  

```

```{r "COMPARE WITH MOCK DATA", eval = FALSE, fig.width = 18, fig.height = 15}

# Plot top features across data
plt_feats <- degs %>%
  group_by(subtype, feature, group) %>%
  summarize(n = n_distinct(mouse), .groups = "drop") %>%
  arrange(desc(n)) %>%
  filter(n > 1) %>%
  filter(group == "high") %>%
  split(.$subtype) %>%
  map(pull, feature)

# plt_feats <- top_degs %>%
#   split(.$group)

clmns <- str_c(names(plt_feats), seq_along(plt_feats))

plt_dat <- lec_so %>%
  Seurat::AddModuleScore(features = list(plt_feats$high$feature)) %>%
  # Seurat::AddModuleScore(features = plt_feats, name = names(plt_feats)) %>%
  FetchData(c("Ag_class_2", "Cluster1", "orig.ident", "mouse", "tm", "subtype"))

plt_dat %>%
  mutate(Ag_class_2 = fct_relevel(Ag_class_2, c("low", "high"))) %>%
  ggplot(aes(mouse, Cluster1, fill = mouse, alpha = Ag_class_2)) +
  geom_boxplot() +
  scale_alpha_manual(values = c(high = 0.75, low = 0.25)) +
  facet_wrap(~ subtype) +
  djvdj_theme()

# Load mock data
chikv_so <- qread("~/Dropbox/Ryan/Projects/morrison-lnsc/results/sobjs/so_merge.qs")

# Calculate module score
chikv_dat <- chikv_so %>%
  Seurat::AddModuleScore(features = plt_feats, name = names(plt_feats)) %>%
  FetchData(c("orig.ident", "lec_type", "tm", "cell_type", "rep", "cell_type_2", "treatment", clmns, "subtype"))

# Calculate mean signal
all_gns <- unique(unlist(plt_feats))

chikv_dat <- chikv_so %>%
  FetchData(c(
    "orig.ident", "lec_type", "tm", "cell_type",
    "rep", "cell_type_2", "treatment", "subtype",
    all_gns
  ))

# Plot example genes
chikv_so %>%
  FetchData(c("treatment", "tm", "lec_type", "rep", "Ctsd", "cell_type_2")) %>%
  filter(cell_type_2 == "LEC" | lec_type == "BEC") %>%
  mutate(
    treat_rep = str_c(treatment, "-", rep),
    treatment = fct_relevel(treatment, c("mock", "CHIKV"))
  ) %>%
  ggplot(aes(lec_type, Ctsd, fill = treatment, alpha = rep)) +
  geom_boxplot(outlier.size = 0.1) +
  scale_alpha_manual(values = c(0.2, 0.4, 0.6)) +
  facet_wrap(~ tm) +
  djvdj_theme() +
  theme(
    aspect.ratio = 0.4,
    axis.title.x = element_blank(),
    axis.text.x  = element_text(angle = 45, hjust = 1)
  )

# # Plot mean list expression
# plt_dat <- chikv_dat %>%
#   pivot_longer(any_of(all_gns)) %>%
#   filter(cell_type_2 == "LEC" | lec_type == "BEC")
# 
#   mutate(
#     treat_rep = str_c(treatment, "-", rep),
#     treatment = fct_relevel(treatment, c("mock", "CHIKV"))
#   ) %>%
#   ggplot(aes(lec_type, Ctsd, fill = treatment, alpha = rep)) +
#   geom_boxplot(outlier.size = 0.1) +
#   scale_alpha_manual(values = c(0.2, 0.4, 0.6)) +
#   facet_wrap(~ tm) +
#   djvdj_theme() +
#   theme(
#     aspect.ratio = 0.4,
#     axis.title.x = element_blank(),
#     axis.text.x  = element_text(angle = 45, hjust = 1)
#   )

# Plot module scores
plt <- clmns %>%
  map(~ {
    chikv_dat %>%
      filter(cell_type_2 == "LEC" | lec_type == "BEC") %>%
      mutate(
        treat_rep = str_c(treatment, "-", rep),
        treatment = fct_relevel(treatment, c("mock", "CHIKV"))
      ) %>%
      ggplot(aes(lec_type, !!sym(.x), fill = treatment, alpha = rep)) +
      geom_boxplot(outlier.size = 0.01, outlier.alpha = 1, size = 0.4) +
      scale_alpha_manual(values = c(0.2, 0.4, 0.6)) +
      facet_wrap(~ tm, nrow = 1) +
      guides(alpha = FALSE) +
      djvdj_theme() +
      theme(
        aspect.ratio = 0.4,
        axis.title.x = element_blank(),
        axis.text.x  = element_text(angle = 45, hjust = 1),
        legend.position = "top"
      )
  }) %>%
  plot_grid(plotlist = ., ncol = 2)

ggsave(
  here(params$so_dir, "ag_markers_chikv.png"),
  plot = plt,
  width = 18, height = 15
)
```



```{r "COMPARISON OF EMPTY DROPS", eval = FALSE}
# Paths to raw matrices to use for normalization factors
mat_dir <- names(params$res_dir)
mat_dir <- set_names(params[mat_dir], mat_dir)

# Calculate normalization factors
filt_counts <- qread(here("results/2022-10-28/sobjs/filt_norm_factors.qs"))
raw_counts  <- qread(here("results/2022-10-28/sobjs/norm_factors.qs"))

counts_df <- list(raw = raw_counts, filt = filt_counts) %>%
  imap_dfr(~ {
    reads <- .y
    
    .x %>%
      imap_dfr(~ {
        tibble(
          reads   = reads,
          run     = .y,
          library = names(.x),
          counts  = .x
        )
      })
  }) %>%
  pivot_wider(names_from = reads, values_from = counts) %>%
  mutate(frac_in_cells = filt / raw) %>%
  arrange(desc(frac_in_cells))

# Sample meta.data
lib_meta <- lec_so@meta.data %>%
  distinct(orig.ident, sample, tm, mouse, exp) %>%
  as_tibble()

counts_df %>%
  left_join(lib_meta, by = c(library = "orig.ident")) %>%
  filter(!is.na(sample)) %>%
  arrange(desc(frac_in_cells))
```

```{r "DSB TEST", eval = FALSE}
# Test the dsb package for correcting CITE-seq data based on empty droplets
# the results are not very convincing
# the signals do not decrease with time and the 3wk signals appear much higher
# than all the other timepoints

# data.frame with parameters
library(dsb)
library(furrr)

mat_dir <- names(params$res_dir)
mat_dir <- set_names(params[mat_dir], mat_dir)

mat_df <- mat_dir %>%
  imap_dfr(~ {
    rn <- .y
    
    .x %>%
      imap_dfr(~ {
        tibble(
          run    = rn,
          sample = .y,
          path   = here(params$res_dir[[rn]], .x, "outs"),
          rep    = str_extract(run, "^r[0-9]+(?=_)")
        )
      })
  })

# Normalize Ag matrices
plan(multisession, workers = 6)

norm_mats <- mat_df %>%
  future_pmap(~ {
    args <- list(...)
    filt_mat <- Read10X(here(args$path, "filtered_feature_bc_matrix"))$`Antibody Capture`
    raw_mat  <- Read10X(here(args$path, "raw_feature_bc_matrix"))$`Antibody Capture`
    raw_mat  <- raw_mat[, !colnames(raw_mat) %in% colnames(filt_mat)]
    
    DSBNormalizeProtein(
      cell_protein_matrix = filt_mat,
      empty_drop_matrix   = raw_mat,
      denoise.counts      = FALSE,
      pseudocount.use     = 1,
      use.isotype.control = FALSE,
      return.stats        = TRUE
    )
  })

names(norm_mats) <- mat_df$sample

# Create data.frame for corrected data
dsb_dat <- mat_df %>%
  pmap_dfr(~ {
    args <- list(...)
    
    dat <- t(norm_mats[[args$sample]]$dsb_normalized_matrix)
    
    if (!grepl("^CD45", args$sample)) {
      colnames(dat) <- params$ag_key[[args$rep]][colnames(dat)]
    }
    
    colnames(dat) <- colnames(dat) %>%
      str_c("dsb_", .)
    
    dat %>%
      as_tibble(rownames = "raw_cell_id") %>%
      mutate(run = args$fun, sample = args$sample)
  })

# Add corrected counts to object
lec_so <- qread(here(params$so_dir, "results/2022-10-28/sobjs/so_lec_new.qs"))

lec_so@meta.data <- lec_so@meta.data %>%
  mutate(raw_cell_id = str_remove(rownames(.), "_[_0-9]+$")) %>%
  as_tibble(rownames = ".cell_id") %>%
  left_join(dsb_dat, by = c("raw_cell_id", orig.ident = "sample")) %>%
  mutate(
    Ag_score = case_when(
      grepl("d[0-9]+$", mouse) ~ dsb_ovalbumin,
      grepl("^3wk$", mouse)    ~ dsb_3wk,
      grepl("^6wk$", mouse)    ~ dsb_6wk,
      
      Ag_3wk_score >= Ag_6wk_score ~ Ag_3wk_score,  # For 6wk-3wk use highest
      Ag_6wk_score >= Ag_3wk_score ~ Ag_6wk_score
    ),
    Ag_score_3 = ifelse(vaccination == "dual", Ag_3wk_score, Ag_score),
    Ag_score_6 = ifelse(vaccination == "dual", Ag_6wk_score, Ag_score),
    tm_3 = ifelse(vaccination == "dual", 21, tm),
    tm_6 = ifelse(vaccination == "dual", 42, tm)
  ) %>%
  column_to_rownames(".cell_id")

# Create plots
# timepoints
lec_so@meta.data %>%
  # filter(subtype == "fLEC") %>%
  # filter(tm != 21) %>%
  filter(vaccination == "single") %>%
  ggplot(aes(tm, Ag_score, group = tm, fill = subtype)) +
  geom_boxplot(width = 3) +
  facet_wrap(~ subtype, nrow = 1) +
  djvdj_theme()

# antigen-exchange
lec_so@meta.data %>%
  # filter(subtype == "fLEC") %>%
  mutate(vaccination = fct_relevel(vaccination, c("single", "dual"))) %>%
  ggplot(aes(tm_3, Ag_score_3, group = tm, fill = subtype, color = subtype, alpha = vaccination)) +
  geom_boxplot(width = 6, position = position_dodge2(preserve = "single")) +
  facet_wrap(~ subtype, nrow = 1) +
  scale_alpha_manual(values = c(0.25, 0.75)) +
  djvdj_theme()

lec_so@meta.data %>%
  # filter(tm_6 != 21) %>%
  # filter(subtype == "fLEC") %>%
  mutate(vaccination = fct_relevel(vaccination, c("single", "dual"))) %>%
  ggplot(aes(tm_6, Ag_score_6, group = tm, fill = subtype, color = subtype, alpha = vaccination)) +
  geom_boxplot(width = 6, position = position_dodge2(preserve = "single")) +
  facet_wrap(~ subtype, nrow = 1) +
  scale_alpha_manual(values = c(0.25, 0.75)) +
  djvdj_theme()
```

```{r "PLOT RAW COUNTS CORRELATION", eval = FALSE}
clmns <- c(
  "exp", "mouse", "vaccination", "orig.ident", "subtype",
  "adt_ABPS2", "adt_ABPS4",
  "Ag_3wk", "Ag_6wk"
)

dat <- lec_so %>%
  FetchData(clmns, slot = "counts") %>%
  filter(vaccination == "dual") %>%
  as_tibble(rownames = "cell_id")

xvar <- "adt_ABPS2"
yvar <- "adt_ABPS4"

# xvar <- "Ag_3wk"
# yvar <- "Ag_6wk"

dat %>%
  ggplot(aes(!!sym(xvar) + 1, !!sym(yvar) + 1, color = subtype)) +
  geom_point() +
  facet_grid(exp ~ subtype) +
  djvdj_theme() +
  theme(aspect.ratio = 1) +
  geom_abline(slope = 1, intercept = 0) +
  scale_x_log10() +
  scale_y_log10()
```