---
title: "GeoMx figures"
author: "Ryan Sheridan"
date: "`r Sys.Date()`"
output: html_document
params:
  template_dir:  "src"                                                     # template Rmarkdowns
  so_dir:        "~/Dropbox/Ryan/Projects/antigen-exchange/results/sobjs"  # Seurat objects
  mod_dir:       "~/Dropbox/Ryan/Projects/antigen-exchange/results/models/2023-11-01" # ML models
  geomx_dir:     "results/geomx"
  chikv_dir:     "~/Dropbox/Ryan/Projects/morrison-lnsc/results/sobjs"     # CHIKV objects
  ref_dir:       "ref"                                                     # clustifyr references
  sample_info:   "sample_info.xlsx"
  res_dir:                                                
    value:
      exp-0: "~/Projects/antigen-tracking/results"
      exp-1: "~/Projects/tamburini-antigen-tracking/results/2022-03-11"
      exp-2: "~/Projects/tamburini-antigen-tracking/results/2022-10-28"
editor_options: 
  chunk_output_type: inline
---

```{r "setup", include = FALSE}
# Default chunk options
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo    = FALSE,
  dpi     = 300
)

# Increase memory for loading image
options(java.parameters = "-Xmx12000m")

knitr::knit(here::here(params$template_dir, "setup.Rmd"), "")
knitr::knit(here::here(params$template_dir, "setup-geomx.Rmd"), "")

rm(lec_so, dc_so)

geomx_dat <- geomx %>%
  pData() %>%
  as_tibble(rownames = "dcc")

# Mouse key
geo_m_key <- c(
  `3wk`     = "day 21",
  `6wk`     = "day 42",
  `6wk-3wk` = "dual",
  naive     = "naive"
)
```

---

<br>

A.  Annotated regions used for antigen quantification are shown for a
    representative lymph node.
B.  Segment type is shown below for a representative lymph node.
C.  Antigen signal is shown below for a representative lymph node.
D.  21 day Ag signal is shown for Lyve1+ and CD11C+ segments from mice that
    received a single 21 day vaccination and mice that received sequential 21/42
    day vaccinations. The 42 day samples were excluded due to the low signal
    observed at this later timepoint. The number of segments plotted is shown
    above each boxplot.
E.  Ag signal is shown for Lyve1+ and CD11C+ segments. Note the strong
    correlation between Ag signal for Lyve1+ and CD11C+ segments.

```{r "cropped images", fig.width = 14, fig.height = 7}
plot_geomx_image <- function(obj_in, color, clrs = NULL, labs = waiver(),
                             legend_values = names(labs) %||% names(clrs),
                             ttl = NULL, line_size = 0.5, scale_size = 8,
                             scale_dist = 1.5, ...) {
  
  res <- obj_in %>%
    plotSpatialOverlay(
      colorBy          = color,
      scaleBar         = TRUE,
      corner           = "bottomleft",
      textDistance     = scale_dist,
      scaleBarLineSize = line_size,
      scaleBarFontSize = scale_size / .pt,
      ...
    ) +
    ggtitle(ttl) +
    theme(
      plot.margin  = margin(r = 15, l = 15),
      plot.title   = element_text(hjust = 0.5, size = ttl_pt2 * 1.5),
      legend.title = element_blank()
    )
  
  if (!is.null(legend_values)) {
    all_sams_df <- tibble(x = 1, y = 1, colorBy = legend_values)
    
    res <- res +
      geom_blank(aes(x, y, fill = colorBy), data = all_sams_df)
  }
  
  if (!is.null(clrs)) {
    if (is.numeric(obj_in@plottingFactors[[color]])) {
      res <- res +
        scale_fill_gradientn(colours = clrs) +
        guides(fill = guide_colorbar(ticks = FALSE)) +
        theme(
          legend.position   = "bottom",
          legend.key.width  = unit(50, "pt"),
          legend.key.height = unit(7, "pt"),
          legend.text       = element_text(size = txt_pt2)
        )
    } else {
      res <- res +
        scale_fill_manual(values = clrs, labels = labs) +
        theme(legend.text = element_text(size = ttl_pt2))
    }
  }
  
  res
}

# Select LN to plot
sams <- geomx_dat %>%
  filter(
    LN == "ova-GeoMX+VV 6week and 3week LN1",
    Segment %in% c("Lyve1", "CD11C")
  ) %>%
  pull(Sample_ID)

img_cropped <- geomx_imgs$`Slide 1` %>%
  cropSamples(sampleIDs = sams, sampsOnly = TRUE)

img_args <- list(
  color = c("Region", "Segment", "Ag-score"),
  clrs  = list(reg_clrs, seg_clrs, c("white", "#D7301F")),
  ttl   = c("Region", "Segment type", "Ag signal")
)

# Plot cropped image
img_fig <- img_args %>%
  pmap(~ {
    args <- list(...)
    
    img_cropped %>%
      plot_geomx_image(
        color = args$color,
        clrs  = args$clrs,
        ttl   = args$ttl,
        scale_size = ttl_pt1,
        scale_dist = 3,
        legend_values = NULL,
      ) +
      theme(legend.position = "bottom")
  }) %>%
  plot_grid(
    plotlist = .,
    nrow  = 1,
    align = "h",
    axis  = "tb",
    labels = c("A", "B", "C"),
    label_size = ttl_pt2 * 1.5,
    hjust = 0.05
  )
```

```{r "geomx Ag boxplots", fig.height = 4}
# Signals to plot
ag_clmns <- c(
  # "Ag_score" = "Ag_score",
  "Ag signal (21 day)" = "Ag_3wk_score",
  "Ag signal (42 day)" = "Ag_6wk_score"
)

# Function for strip labels
geo_lab_fn <- function(x) {
  key <- c(
    "3wk"     = "21 day",
    "6wk"     = "42 day",
    "6wk-3wk" = "dual"
  )
  
  idx <- x %in% names(key)
  
  x[idx]  <- key[x[idx]]
  x[!idx] <- str_c(x[!idx], "+")
  
  x
}

# Create boxplot
geo_bxs <- ag_clmns[1] %>%
  imap(~ {
    dat <- geomx_dat %>%
      # filter(sample != "naive", Segment %in% c("Lyve1", "CD11C")) %>%
      filter(
        sample %in% c("3wk", "6wk-3wk"),
        Segment %in% c("Lyve1", "CD11C")
      ) %>%
      mutate(Segment = fct_relevel(Segment, names(seg_clrs)))
    
    n_dat <- dat %>%
      group_by(region_3, Segment) %>%
      summarize(n = n_distinct(dcc), .groups = "drop")
    
    dat %>%  
      ggplot(aes(region_3, !!sym(.x), fill = region_3)) +
      geom_boxplot(key_glyph = draw_key_point, alpha = 1, outlier.size = 0.25, width = 0.65) +
      geom_text(
        aes(region_3, Inf, label = n, fill = NULL),
        data  = n_dat,
        vjust = 1.4,
        size  = 10 / .pt
      ) +
      
      facet_grid(~ Segment, scales = "free_y", labeller = as_labeller(geo_lab_fn)) +
      scale_fill_manual(values = reg_clrs) +
      scale_y_continuous(expand = expansion(c(0.05, 0.2))) +
      labs(y = .y) +
      base_theme +
      theme(
        aspect.ratio    = 1.4,
        legend.position = "none",
        axis.text.x     = element_text(size = ttl_pt2, angle = 45, hjust = 1),
        axis.title.y    = element_text(size = ttl_pt2),
        axis.title.x    = element_blank()
      )
  })
```

```{r "geomx Ag scatter"}
# Scatter plot comparing Lyve1 and CD11C segments
clmn <- "Ag_score"

dat <- geomx_dat %>%
  filter(
    Segment %in% c("Lyve1", "CD11C"),
    sample != "naive"
  ) %>%
  dplyr::select(
    `ROI Coordinate X`, `ROI Coordinate Y`,
    region_3, Segment, sample,
    !!sym(clmn)
  ) %>%
  group_by(`ROI Coordinate X`, `ROI Coordinate Y`) %>%
  filter(all(c("Lyve1", "CD11C") %in% Segment)) %>%
  ungroup() %>%
  mutate(Segment = str_c(Segment, "+ Ag signal")) %>%
  pivot_wider(names_from = Segment, values_from = !!sym(clmn))

# Correlation
cor <- cor(dat$`CD11C+ Ag signal`, dat$`Lyve1+ Ag signal`, method = "pearson")
cor <- round(cor, digits = 2)
cor = str_c("italic(r) == ", cor)

sct_clrs <- reg_clrs
sct_clrs["cortex"] <- "black"

geo_sct <- dat %>%
  ggplot(aes(`CD11C+ Ag signal`, `Lyve1+ Ag signal`, color = region_3)) +
  geom_smooth(method = "lm", color = "black", alpha = 0.1, linewidth = 0.5, linetype = 2) +
  geom_point(size = 2) +
  geom_text(
    aes(x = -Inf, y = Inf, label = cor),
    parse = TRUE,
    hjust = -0.3,
    vjust = 1.4,
    color = "black",
    size  = ttl_pt1 / .pt,
    check_overlap = TRUE
  ) +
  scale_color_manual(values = sct_clrs) +
  base_theme +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.text  = element_text(size = ttl_pt2),
    axis.title   = element_text(size = ttl_pt2)
  )
```

```{r "geomx figure", fig.width = 14, fig.height = 10}
bot <- plot_grid(
  geo_bxs[[1]], geo_sct,
  nrow   = 1,
  align  = "h",
  axis   = "tb",
  labels = c("D", "E"),
  label_size = ttl_pt2 * 1.5
)

plot_grid(
  img_fig, bot,
  ncol = 1,
  rel_heights = c(1, 0.75)
)

```

<br>

## Extended figures

The sample identity is shown below for each slide.

```{r "all geomx samples", fig.width = 10, fig.height = 8}
geomx_imgs %>%
  imap(~ {
    .x %>%
      plot_geomx_image(
        color = "Sample",
        clrs  = geo_m_clrs,
        labs  = geo_m_key,
        ttl   = .y
      )
  }) %>%
  wrap_plots(
    guides = "collect",
    nrow = 1
  ) &
  theme(legend.position = "bottom")
```

<br>

Revised region labels are shown below for each slide.

```{r "all geomx regions", fig.width = 10, fig.height = 8}
geomx_imgs %>%
  imap(~ {
    .x %>%
      plot_geomx_image(
        color = "Region",
        clrs  = reg_clrs,
        ttl   = .y
      )
  }) %>%
  wrap_plots(
    guides = "collect",
    nrow = 1
  ) &
  theme(legend.position = "bottom")
```

<br>

Segment labels are shown below for each slide.

```{r "all geomx segments", fig.width = 10, fig.height = 8}
geomx_imgs %>%
  imap(~ {
    .x %>%
      plot_geomx_image(
        color = "Segment",
        clrs  = seg_clrs,
        ttl   = .y
      )
  }) %>%
  wrap_plots(
    guides = "collect",
    nrow = 1
  ) &
  theme(legend.position = "bottom")
```

<br>

Ag signal is shown for single and dual vaccinated mice for each region and
segment type (e.g. Lyve1+). The number of segments included is shown above each
boxplot.

```{r "all geomx boxplots Ag-score", fig.width = 10, fig.height = 6.5}
# Create boxplot
ag_clmns %>%
  imap(~ {
    dat <- geomx_dat %>%
      filter(sample != "naive", Segment != "Other") %>%
      mutate(Segment = fct_relevel(Segment, names(seg_clrs)))
    
    n_dat <- dat %>%
      group_by(region_3, Segment, sample) %>%
      summarize(n = n_distinct(dcc), .groups = "drop")
    
    dat %>%  
      ggplot(aes(region_3, !!sym(.x), fill = region_3)) +
      geom_boxplot(key_glyph = draw_key_point, alpha = 1, outlier.size = 0.25, width = 0.65) +
      geom_text(
        aes(region_3, Inf, label = n, fill = NULL),
        data  = n_dat,
        vjust = 1.4,
        size  = 8 / .pt
      ) +
      
      facet_grid(Segment ~ sample, scales = "free_y", labeller = as_labeller(geo_lab_fn)) +
      scale_fill_manual(values = reg_clrs) +
      scale_y_continuous(expand = expansion(c(0.05, 0.2))) +
      labs(y = .y) +
      base_theme +
      theme(
        aspect.ratio    = 1.4,
        legend.position = "none",
        axis.text.x     = element_text(angle = 45, hjust = 1),
        axis.title.x    = element_blank()
      )
  }) %>%
  plot_grid(
    plotlist = .,
    nrow  = 1,
    align = "h",
    axis  = "tb"
  )
```



```{r "all geomx Ag-score", fig.width = 10, fig.height = 8, eval = FALSE}
geomx_imgs %>%
  imap(~ {
    .x %>%
      plot_geomx_image(
        color = "Ag-score",
        clrs  = c("white", "#D7301F", "#D7301F"),
        ttl   = .y
      ) +
      guides(fill = guide_colorbar(ticks = FALSE))
  }) %>%
  wrap_plots(nrow = 1) &
  theme(
    legend.position   = "bottom",
    legend.key.width  = unit(40, "pt"),
    legend.key.height = unit(7, "pt"),
    legend.text       = element_text(size = 8),
    legend.title      = element_text(size = txt_pt1)
  )
```

```{r "EACH SEPARATE", eval = FALSE}
var <- "Region"

geomx_imgs[1] %>%
  map(~ {
    obj  <- .x
    sams <- obj@plottingFactors$Sample %>%
      as.character() %>%
      unique() %>%
      na.omit()

    sams %>%
      map(~ {
        ids <- obj@plottingFactors %>%
          dplyr::filter(Sample == .x) %>%
          rownames() %>%
          unique()
        
        img <- obj %>%
          cropSamples(
            sampleIDs = ids, sampsOnly = FALSE, buffer = 0.2
          )

        img %>%
          plotSpatialOverlay(
            colorBy      = var,
            scaleBar     = TRUE,
            corner       = "bottomleft",
            textDistance = 5
          ) +
          ggtitle(.x) +
          geomx_theme
      }) %>%
      plot_grid(
        plotlist = .,
        ncol = 1
      )
  })
```

```{r "CROPPED IMAGES", eval = FALSE}
# Crop image to show selected ROIs
rois <- c("=\"006\"", "=\"007\"", "=\"008\"", "=\"022\"")

sams <- annots %>%
  filter(
    roi %in% rois,
    segment %in% c("Lyve1", "CD11C")
  ) %>%
  pull(Sample_ID)

# img_cropped <- img %>%
#   cropSamples(sampleIDs = sams, sampsOnly = TRUE)
```

```{r, eval = FALSE}
# Plot cropped image
# fig_dir <- "~/Dropbox/Ryan/Documents/Hesselberth_lab/Meetings/2024-03-12_data_club"

geomx_theme <- theme(
  legend.position = "bottom",
  legend.text     = element_text(size = 24),
  legend.title    = element_blank()
)

# img_cropped %>%
#   plotSpatialOverlay(
#     colorBy      = "Segment",
#     scaleBar     = TRUE,
#     corner       = "bottomleft",
#     textDistance = 5
#   ) +
#   scale_fill_manual(values = unname(seg_clrs)) +
#   geomx_theme
# 
# ggsave(here(fig_dir, "geomx_img_fig_1.png"), dpi = 400)

img %>%
  plotSpatialOverlay(
    colorBy      = "Sample",
    scaleBar     = TRUE,
    corner       = "bottomleft",
    textDistance = 5
  ) +
  # scale_fill_manual(values = seg_clrs) +
  geomx_theme

ggsave(here(fig_dir, "geomx_img_fig_2.png"), dpi = 400)

img_cropped %>%
  plotSpatialOverlay(
    colorBy      = "Ova",
    scaleBar     = TRUE,
    corner       = "bottomleft",
    textDistance = 5
  ) +
  scale_fill_gradient2(
    low = "grey", high = "#D7301F", mid = "#F0E442",
    midpoint = 2500
  )

ggsave(here(fig_dir, "geomx_img_fig_3.png"), dpi = 400)
```


